---
title: "02 Mechanistic Model"
author: "Jianqiu Zheng"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
source("R/helpers.R")
```

## Purpose

This notebook follows the mechanistic-model section of the original `S19S_sed_model_graph.Rmd` workflow. It does two things:

1. Recreates the theoretical perturbation plots used for the original Figure 2.
2. Recreates the measured mechanistic scenarios used for the original Figure 3.

It also keeps the downstream output `derived/mechanistic_predictions.csv`, which is used by the KGML residual-learning workflows.

## Theoretical Mechanistic Perturbations

```{r theoretical-perturbations}
set.seed(10)

Vmax <- 1.0
Cb <- 1.0
y <- 1.0
p <- 1.0

inv_oc <- seq(0, 5, length.out = 500)

df_the <- tibble(
  inv_OC = inv_oc,
  R = Vmax * Cb * exp(-(y / p) * inv_oc)
)

p_the_npoc <- ggplot(df_the, aes(x = inv_OC, y = R)) +
  geom_point(shape = 1, size = 1, color = puor$purple) +
  labs(x = "1/NPOC", y = expression(R[pred])) +
  theme_review()

df_y <- tibble(
  inv_OC = inv_oc,
  y_abs = runif(500, min = 0.1, max = 1),
  R = Vmax * Cb * exp(-(y_abs / p) * inv_oc)
)

p_the_y_curve <- ggplot() +
  geom_point(data = df_y, aes(x = inv_OC, y = R),
             shape = 1, size = 1, color = puor$orange) +
  geom_line(data = df_the, aes(x = inv_OC, y = R),
            linewidth = 0.8, color = puor$purple) +
  labs(x = "1/NPOC", y = expression(R[pred])) +
  theme_review()

p_the_y <- ggplot(df_y, aes(x = y_abs, y = R)) +
  geom_point(shape = 1, size = 1, color = puor$orange) +
  labs(x = "|yi|", y = expression(R[pred])) +
  theme_review()

df_cb <- tibble(
  inv_OC = inv_oc,
  Cb = rnorm(500, mean = 1, sd = 0.25),
  R = Vmax * Cb * exp(-(1 / p) * inv_oc)
)

p_the_cb_curve <- ggplot() +
  geom_point(data = df_cb, aes(x = inv_OC, y = R),
             shape = 1, size = 1, color = puor$orange) +
  geom_line(data = df_the, aes(x = inv_OC, y = R),
            linewidth = 0.8, color = puor$purple) +
  labs(x = "1/NPOC", y = expression(R[pred])) +
  theme_review()

p_the_cb <- ggplot(df_cb, aes(x = Cb, y = R)) +
  geom_point(shape = 1, size = 1, color = puor$orange) +
  labs(x = "Cb", y = expression(R[pred])) +
  theme_review()

df_vh <- tibble(
  inv_OC = inv_oc,
  Vh = runif(500, min = 0.1, max = 1),
  R = rescale01(Vmax * exp(-(1 / Vh) * inv_oc))
)

p_the_vh_curve <- ggplot() +
  geom_point(data = df_vh, aes(x = inv_OC, y = R),
             shape = 1, size = 1, color = puor$orange) +
  geom_line(data = df_the, aes(x = inv_OC, y = R),
            linewidth = 0.8, color = puor$purple) +
  labs(x = "1/NPOC", y = expression(R[pred])) +
  theme_review()

p_the_vh <- ggplot(df_vh, aes(x = Vh, y = R)) +
  geom_point(shape = 1, size = 1, color = puor$orange) +
  labs(x = "Vh", y = expression(R[pred])) +
  theme_review()

ggsave(figure_file("mechanistic_theory_npoc.pdf"), p_the_npoc,
       width = 3.5, height = 3)
ggsave(figure_file("mechanistic_theory_y.pdf"), p_the_y,
       width = 3.5, height = 3)
ggsave(figure_file("mechanistic_theory_cb.pdf"), p_the_cb,
       width = 3.5, height = 3)
ggsave(figure_file("mechanistic_theory_vh.pdf"), p_the_vh,
       width = 3.5, height = 3)

if (requireNamespace("patchwork", quietly = TRUE)) {
  p_theory_combined <- (p_the_npoc | p_the_y_curve) /
    (p_the_cb_curve | p_the_vh_curve)

  ggsave(figure_file("mechanistic_revised_figure2_theory.pdf"),
         p_theory_combined, width = 7, height = 5.6)
  p_theory_combined
} else {
  p_the_npoc
  p_the_y_curve
  p_the_cb_curve
  p_the_vh_curve
}
```

## Load Peak-Level Modeling Inputs

```{r load-peak-level-data}
process_site <- function(path, position, filter_lambda = FALSE) {
  read_csv(path, col_types = cols(site = col_character())) %>%
    mutate(
      Position = position,
      factor = (water_mass + 10) * 0.01 / sed_mass,
      respC = rate * factor * (12 / 32),
      npoc = conc * factor
    ) %>%
    { if (filter_lambda) filter(., lambda > 0) else . } %>%
    select(-factor)
}

up_v <- process_site(derived_file("up_all.csv"), "U", filter_lambda = TRUE)
mid_v <- process_site(derived_file("mid_all.csv"), "M", filter_lambda = TRUE)
low_v <- process_site(derived_file("low_all.csv"), "D", filter_lambda = TRUE)

all_data <- bind_rows(
  up_v %>% mutate(level = "up"),
  mid_v %>% mutate(level = "mid"),
  low_v %>% mutate(level = "low")
)

all_data %>%
  summarise(
    peak_rows = n(),
    sites = n_distinct(site),
    site_positions = n_distinct(paste(site, Position)),
    complete_rows = sum(complete.cases(rate, conc, biomass, sed_mass,
                                       water_mass, npoc, respC, ssa_tot,
                                       stoichMet_donor, ne, lambda))
  )
```

## Scenario 1: FTICR Thermodynamics and NPOC

```{r scenario-1-thermo-npoc}
summary_all <- all_data %>%
  drop_na() %>%
  mutate(
    pr1 = exp(-abs(stoichMet_donor) / npoc),
    pr2 = (3 / ne) * exp(-abs(stoichMet_donor) / npoc)
  ) %>%
  group_by(site, Position) %>%
  summarise(
    across(where(is.numeric) & -any_of(c("pr1", "pr2")),
           ~ mean(.x, na.rm = TRUE), .names = "{.col}_mean"),
    across(c(pr1, pr2),
           list(p05 = ~ quantile(.x, 0.05, na.rm = TRUE),
                p50 = ~ quantile(.x, 0.50, na.rm = TRUE),
                p95 = ~ quantile(.x, 0.95, na.rm = TRUE)),
           .names = "{.col}_{.fn}"),
    .groups = "drop"
  ) %>%
  select(site, Position, everything())

write_csv(summary_all, derived_file("mechanistic_original_summary_all_levels.csv"))

fticr_features <- read_csv(raw_file("fticr_features.csv"),
                           col_types = cols(site = col_character()))

class_cols <- c(
  "frac_ConHC", "frac_Lipid", "frac_AminoSugar", "frac_Protein",
  "frac_Tannin", "frac_UnsatHC", "frac_Lignin", "frac_Carb", "frac_Other"
)

fticr_site_composition <- fticr_features %>%
  group_by(site) %>%
  summarise(
    n_site_positions = n_distinct(Position),
    n_peaks = mean(n_peaks, na.rm = TRUE),
    across(all_of(class_cols), ~ mean(.x, na.rm = TRUE)),
    .groups = "drop"
  )

site_order <- fticr_site_composition %>%
  arrange(as.integer(site)) %>%
  pull(site)

fticr_site_composition <- fticr_site_composition %>%
  mutate(site = factor(site, levels = site_order))

fticr_class_long <- fticr_site_composition %>%
  select(site, all_of(class_cols)) %>%
  pivot_longer(
    cols = -site,
    names_to = "Compound_class",
    values_to = "Fraction"
  ) %>%
  mutate(
    Compound_class = str_remove(Compound_class, "^frac_"),
    Compound_class = factor(
      Compound_class,
      levels = str_remove(class_cols, "^frac_"),
      labels = c("Condensed HC", "Lipid", "Amino sugar", "Protein",
                 "Tannin", "Unsat. HC", "Lignin", "Carbohydrate", "Other")
    )
  )

write_csv(fticr_site_composition,
          derived_file("supplement_fticr_site_composition_summary.csv"))
write_csv(fticr_class_long,
          derived_file("supplement_fticr_site_compound_classes.csv"))

p_peak_richness <- ggplot(fticr_site_composition,
                          aes(x = n_peaks, y = site)) +
  geom_point(color = puor$purple, size = 1.4) +
  labs(x = "Number\nof peaks", y = "Site") +
  theme_review() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 4),
    axis.ticks.y = element_blank(),
    panel.grid.major.x = element_line(color = "grey85", linewidth = 0.25)
  )

p_class_composition <- ggplot(fticr_class_long,
                              aes(x = Fraction, y = site, fill = Compound_class)) +
  geom_col(width = 0.85, color = "white", linewidth = 0.08) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.01))) +
  scale_fill_manual(values = puor_seq(nlevels(fticr_class_long$Compound_class)),
                    name = "Compound class") +
  labs(x = "Compound-class composition", y = NULL) +
  theme_review() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right"
  )

if (requireNamespace("patchwork", quietly = TRUE)) {
  p_om_variation <- p_peak_richness + p_class_composition +
    patchwork::plot_layout(widths = c(1.1, 3.2))
} else {
  p_om_variation <- p_class_composition
}

ggsave(figure_file("supplement_fticr_inferred_om_site_variation.pdf"),
       p_om_variation, width = 8, height = 8.5)

p_om_variation

pr1_cols <- c("pr1_p05", "pr1_p50", "pr1_p95")
pr2_cols <- c("pr2_p05", "pr2_p50", "pr2_p95")
rng1 <- range(unlist(summary_all[pr1_cols]), na.rm = TRUE)
rng2 <- range(unlist(summary_all[pr2_cols]), na.rm = TRUE)
span1 <- diff(rng1)
span2 <- diff(rng2)

df_scaled <- summary_all %>%
  mutate(
    across(all_of(pr1_cols),
           ~ if (span1 == 0) 0 else (.x - rng1[1]) / span1,
           .names = "{.col}_scaled"),
    across(all_of(pr2_cols),
           ~ if (span2 == 0) 0 else (.x - rng2[1]) / span2,
           .names = "{.col}_scaled"),
    inv_npoc = 1 / npoc_mean
  )

comp_df <- df_scaled %>%
  mutate(
    res_scaled = (respC_mean - min(respC_mean, na.rm = TRUE)) /
      (max(respC_mean, na.rm = TRUE) - min(respC_mean, na.rm = TRUE))
  ) %>%
  select(site, Position, inv_npoc, res_scaled) %>%
  drop_na()

pr_for_plot <- df_scaled %>%
  select(site, inv_npoc, matches("^pr[12]_(p05|p50|p95)_scaled$")) %>%
  pivot_longer(
    cols = -c(site, inv_npoc),
    names_to = c("variable", "stat"),
    names_pattern = "(pr[12])_(p\\d{2})_scaled",
    values_to = "value"
  ) %>%
  pivot_wider(
    id_cols = c(site, inv_npoc, variable),
    names_from = stat,
    values_from = value
  ) %>%
  mutate(variable = factor(variable, levels = c("pr1", "pr2")))

p_scen1 <- ggplot() +
  geom_ribbon(
    data = pr_for_plot,
    aes(x = inv_npoc, ymin = p05, ymax = p95, fill = variable),
    alpha = 0.3
  ) +
  geom_point(
    data = pr_for_plot,
    aes(x = inv_npoc, y = p50, color = variable, shape = "p50"),
    size = 2, alpha = 0.9
  ) +
  geom_smooth(
    data = pr_for_plot,
    aes(x = inv_npoc, y = p50, color = variable),
    method = "loess", se = FALSE, linewidth = 0.7, show.legend = FALSE
  ) +
  geom_point(
    data = comp_df,
    aes(x = inv_npoc, y = res_scaled, color = "Observed", shape = "Observed"),
    size = 2, alpha = 0.9
  ) +
  scale_fill_manual(
    values = c(pr1 = puor$dark_orange, pr2 = puor$orange),
    name = "Predicted"
  ) +
  scale_color_manual(
    values = c(pr1 = puor$dark_orange, pr2 = puor$orange,
               Observed = puor$light_purple),
    breaks = c("pr1", "pr2", "Observed"),
    labels = c("Chemistry + NPOC", "Chemistry + NPOC + uptake", "Observed"),
    name = "Series"
  ) +
  scale_shape_manual(
    values = c(p50 = 16, Observed = 16),
    labels = c("Predicted p50", "Observed"),
    name = "Points"
  ) +
  labs(x = "1/NPOC",
       y = expression(italic(R)[scaled]~"(Observed & Predicted)")) +
  theme_review() +
  theme(legend.position = "right")

ggsave(figure_file("mechanistic_scenario1_thermo_npoc.pdf"),
       p_scen1, width = 5, height = 3)

mechanistic_scenario1_stats <- bind_rows(
  model_stats(comp_df$res_scaled,
              df_scaled %>% select(site, Position, pr1_p50_scaled) %>%
                left_join(comp_df, by = c("site", "Position")) %>%
                pull(pr1_p50_scaled)) %>%
    mutate(Model = "Mechanistic_pr1"),
  model_stats(comp_df$res_scaled,
              df_scaled %>% select(site, Position, pr2_p50_scaled) %>%
                left_join(comp_df, by = c("site", "Position")) %>%
                pull(pr2_p50_scaled)) %>%
    mutate(Model = "Mechanistic_pr2")
) %>%
  select(Model, everything())

write_csv(mechanistic_scenario1_stats,
          derived_file("mechanistic_scenario1_stats.csv"))

p_scen1
mechanistic_scenario1_stats
```

## Scenario 2: Add Microbial Biomass

```{r scenario-2-biomass}
df_biomass <- df_scaled %>%
  mutate(
    pr3 = exp(-abs(1) / npoc_mean) * biomass_mean,
    pr4 = pr2_p05 * biomass_mean,
    pr5 = pr2_p50 * biomass_mean,
    pr6 = pr2_p95 * biomass_mean
  ) %>%
  drop_na()

rng456 <- range(c(df_biomass$pr4, df_biomass$pr5, df_biomass$pr6),
                na.rm = TRUE)
span456 <- diff(rng456)

df_biomass <- df_biomass %>%
  mutate(
    pr4_scaled = if (span456 == 0) 0 else (pr4 - rng456[1]) / span456,
    pr5_scaled = if (span456 == 0) 0 else (pr5 - rng456[1]) / span456,
    pr6_scaled = if (span456 == 0) 0 else (pr6 - rng456[1]) / span456,
    pr3_scaled = rescale01(pr3),
    res_scaled = rescale01(respC_mean)
  ) %>%
  select(site, Position, inv_npoc, res_scaled, pr3_scaled,
         pr4_scaled, pr5_scaled, pr6_scaled, biomass_mean) %>%
  rename(Obv = res_scaled, Pred3 = pr3_scaled, Pred4 = pr4_scaled,
         Pred5 = pr5_scaled, Pred6 = pr6_scaled, biomass = biomass_mean)

p_scen2 <- ggplot() +
  geom_ribbon(
    data = df_biomass,
    aes(x = inv_npoc, ymin = Pred4, ymax = Pred6, fill = "Pred p05-p95"),
    alpha = 0.3
  ) +
  geom_point(
    data = df_biomass,
    aes(x = inv_npoc, y = Pred5, color = "Pred p50"),
    size = 2, alpha = 0.8
  ) +
  geom_point(
    data = df_biomass,
    aes(x = inv_npoc, y = Pred3, color = "Biomass + NPOC"),
    size = 2, shape = 16
  ) +
  geom_point(
    data = df_biomass,
    aes(x = inv_npoc, y = Obv, color = "Observed"),
    size = 2, alpha = 0.8, shape = 16
  ) +
  scale_fill_manual(values = c("Pred p05-p95" = puor$light_orange),
                    name = NULL) +
  scale_color_manual(
    values = c("Pred p50" = puor$orange,
               "Biomass + NPOC" = puor$dark_orange,
               "Observed" = puor$light_purple),
    name = NULL
  ) +
  labs(x = "1/NPOC",
       y = expression(italic(R)[scaled]~"(Observed & Predicted)")) +
  theme_review() +
  theme(legend.position = "right")

ggsave(figure_file("mechanistic_scenario2_biomass.pdf"),
       p_scen2, width = 5.2, height = 3)

mechanistic_scenario2_stats <- bind_rows(
  model_stats(df_biomass$Obv, df_biomass$Pred3) %>% mutate(Model = "Pred3"),
  model_stats(df_biomass$Obv, df_biomass$Pred4) %>% mutate(Model = "Pred4"),
  model_stats(df_biomass$Obv, df_biomass$Pred5) %>% mutate(Model = "Pred5"),
  model_stats(df_biomass$Obv, df_biomass$Pred6) %>% mutate(Model = "Pred6")
) %>%
  select(Model, everything())

write_csv(mechanistic_scenario2_stats,
          derived_file("mechanistic_scenario2_stats.csv"))

p_scen2
mechanistic_scenario2_stats
```

## Scenario 3: Add Fixed Accessibility to Chemistry

```{r scenario-3-fixed-accessibility}
summarise_fixed_vh <- function(dat, vh_values, prefix = "pr") {
  dat %>%
    drop_na() %>%
    mutate(
      pr7 = (3 / ne) * exp(-abs(stoichMet_donor) / (vh_values[1] * npoc)),
      pr8 = (3 / ne) * exp(-abs(stoichMet_donor) / (vh_values[2] * npoc)),
      pr9 = (3 / ne) * exp(-abs(stoichMet_donor) / (vh_values[3] * npoc))
    ) %>%
    group_by(site, Position) %>%
    summarise(
      across(where(is.numeric) & -any_of(c("pr7", "pr8", "pr9")),
             ~ mean(.x, na.rm = TRUE), .names = "{.col}_mean"),
      across(c(pr7, pr8, pr9),
             list(q05 = ~ quantile(.x, 0.05, na.rm = TRUE),
                  q50 = ~ quantile(.x, 0.50, na.rm = TRUE),
                  q95 = ~ quantile(.x, 0.95, na.rm = TRUE)),
             .names = "{.col}_{.fn}"),
      .groups = "drop"
    ) %>%
    mutate(inv_npoc = 1 / npoc_mean,
           res_scaled = rescale01(respC_mean)) %>%
    select(site, Position, everything())
}

make_fixed_vh_plot <- function(df_processed, labels, file_name) {
  x_col <- "inv_npoc"

  quant_long <- df_processed %>%
    select(site, Position, all_of(x_col),
           matches("^pr7_q(05|50|95)$"),
           matches("^pr8_q(05|50|95)$"),
           matches("^pr9_q(05|50|95)$")) %>%
    pivot_longer(
      cols = matches("^pr[789]_q(05|50|95)$"),
      names_to = c("metric", "stat"),
      names_pattern = "(pr[789])_q(05|50|95)",
      values_to = "value"
    ) %>%
    mutate(metric = recode(metric, pr7 = labels[1],
                           pr8 = labels[2], pr9 = labels[3])) %>%
    pivot_wider(names_from = stat, values_from = value) %>%
    mutate(metric = factor(metric, levels = labels))

  quant_scaled <- quant_long %>%
    group_by(metric) %>%
    mutate(
      qmin = min(`05`, `50`, `95`, na.rm = TRUE),
      qmax = max(`05`, `50`, `95`, na.rm = TRUE),
      `05` = (`05` - qmin) / (qmax - qmin),
      `50` = (`50` - qmin) / (qmax - qmin),
      `95` = (`95` - qmin) / (qmax - qmin)
    ) %>%
    ungroup() %>%
    select(-qmin, -qmax)

  obs_df <- df_processed %>%
    select(site, Position, all_of(x_col), res_scaled) %>%
    rename(res = res_scaled)

  color_values <- setNames(c(puor$dark_orange, puor$orange,
                             puor$light_orange), labels)
  color_values <- c(color_values, Observed = puor$light_purple)

  p <- ggplot(quant_scaled, aes(x = .data[[x_col]])) +
    geom_ribbon(aes(ymin = `05`, ymax = `95`, fill = metric),
                alpha = 0.35, color = NA) +
    geom_point(aes(y = `50`, color = metric), size = 2, alpha = 0.9) +
    geom_smooth(aes(y = `50`, color = metric),
                method = "loess", se = FALSE, linewidth = 1,
                linetype = "solid") +
    geom_point(data = obs_df,
               aes(x = .data[[x_col]], y = res, color = "Observed"),
               inherit.aes = FALSE, size = 2, alpha = 0.9, shape = 16) +
    scale_color_manual(values = color_values, name = NULL) +
    scale_fill_manual(values = color_values[labels], name = NULL) +
    labs(x = "1/NPOC",
         y = expression(italic(R)[scaled]~"(Observed & Predicted)")) +
    theme_review() +
    theme(legend.position = "right")

  ggsave(figure_file(file_name), p, width = 4.8, height = 3)

  list(plot = p, quant_scaled = quant_scaled, obs_df = obs_df)
}

df_fixed_vh <- summarise_fixed_vh(all_data, c(0.05, 0.25, 0.50))
fixed_vh_plot <- make_fixed_vh_plot(
  df_fixed_vh,
  labels = c("Vh = 0.05", "Vh = 0.25", "Vh = 0.50"),
  file_name = "mechanistic_scenario3_chemistry_accessibility.pdf"
)

df_accessibility_only <- all_data %>%
  drop_na() %>%
  mutate(
    pr7 = exp(-1 / (0.5 * npoc)),
    pr8 = exp(-1 / (1.0 * npoc)),
    pr9 = exp(-1 / (2.0 * npoc))
  ) %>%
  group_by(site, Position) %>%
  summarise(
    across(where(is.numeric) & -any_of(c("pr7", "pr8", "pr9")),
           ~ mean(.x, na.rm = TRUE), .names = "{.col}_mean"),
    across(c(pr7, pr8, pr9),
           list(q05 = ~ quantile(.x, 0.05, na.rm = TRUE),
                q50 = ~ quantile(.x, 0.50, na.rm = TRUE),
                q95 = ~ quantile(.x, 0.95, na.rm = TRUE)),
           .names = "{.col}_{.fn}"),
    .groups = "drop"
  ) %>%
  mutate(inv_npoc = 1 / npoc_mean,
         res_scaled = rescale01(respC_mean)) %>%
  select(site, Position, everything())

accessibility_only_plot <- make_fixed_vh_plot(
  df_accessibility_only,
  labels = c("Vh = 0.5", "Vh = 1", "Vh = 2"),
  file_name = "mechanistic_scenario4_accessibility_only.pdf"
)

if (requireNamespace("patchwork", quietly = TRUE)) {
  p_mech_scenarios <- (p_scen1 | p_scen2) /
    (fixed_vh_plot$plot | accessibility_only_plot$plot)

  ggsave(figure_file("mechanistic_revised_figure3_scenarios.pdf"),
         p_mech_scenarios, width = 9.5, height = 6)
  p_mech_scenarios
} else {
  p_scen1
  p_scen2
  fixed_vh_plot$plot
  accessibility_only_plot$plot
}
```

## Fitted Effective `Vh`

```{r fit-effective-vh}
analysis_table <- read_csv(derived_file("s19s_sed_analysis_table.csv"),
                           col_types = cols(site = col_character()))

core <- analysis_table %>%
  filter(complete.cases(respC, npoc, biomass_log10, biomass_raw, ssa_tot,
                        stoichMet_donor_mean, ne_mean, d50_mm),
         respC > 0, npoc > 0, biomass_raw > 0, ssa_tot > 0)

df_vh <- core %>%
  mutate(
    mu_max = 1,
    biomass_for_vh = biomass_log10,
    ratio = respC / (mu_max * biomass_for_vh),
    Vh = if_else(
      !is.finite(ratio) | ratio <= 0 | ratio >= 1,
      NA_real_,
      -(abs(stoichMet_donor_mean) * inv_npoc) / log(ratio)
    ),
    log_Vh = log10(Vh)
  )

Vh0_global <- median(df_vh$Vh, na.rm = TRUE)
Vh0_global
```

```{r mechanistic-baseline}
mechanistic_predictions <- df_vh %>%
  mutate(
    Vh,
    log_Vh,
    Vh0_global = Vh0_global,
    R_mech_global_vh =
      exp(-abs(stoichMet_donor_mean) / (Vh0_global * npoc)) * biomass_log10,
    logR_obs = safe_log10(respC),
    logR_mech_global_vh = safe_log10(R_mech_global_vh),
    dR_global_vh = logR_obs - logR_mech_global_vh
  )

write_csv(mechanistic_predictions, derived_file("mechanistic_predictions.csv"))

model_stats(mechanistic_predictions$logR_obs,
            mechanistic_predictions$logR_mech_global_vh) %>%
  mutate(Model = "Mechanistic global Vh") %>%
  select(Model, everything())
```

```{r vh-diagnostics}
df_vh_diagnostics <- mechanistic_predictions %>%
  filter(is.finite(Vh), Vh > 0) %>%
  mutate(
    biomass_over_npoc = biomass_log10 / npoc,
    npoc_over_ssa = npoc / ssa_tot
  )

cor_label <- function(x, y) {
  idx <- complete.cases(x, y)
  if (sum(idx) < 3) return("")
  ct <- suppressWarnings(cor.test(x[idx], y[idx]))
  paste0(
    "R^2 = ", formatC(unname(ct$estimate)^2, digits = 2, format = "f"),
    "\nP = ", format.pval(ct$p.value, digits = 2, eps = 0.001)
  )
}

p_vh <- ggplot(df_vh_diagnostics, aes(x = Vh)) +
  geom_histogram(aes(y = after_stat(density)),
                 binwidth = 0.05,
                 colour = "black", fill = "white") +
  geom_density(alpha = 0.2, fill = "#e08214") +
  scale_x_log10(
    name = "Vh (effective harvest volume)",
    breaks = c(1e-2, 1e-1, 1e0),
    labels = c(expression(10^-2), expression(10^-1),
               expression(10^0)),
    limits = c(1e-2, 1)
  ) +
  geom_vline(
    aes(xintercept = mean(Vh, na.rm = TRUE)),
    color = "#e08214", linetype = "dashed", linewidth = 1
  ) +
  labs(y = "Density") +
  ggpubr::theme_pubr(border = TRUE) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 14, face = "bold")
  )

p_vh_biomass_npoc <- ggplot(df_vh_diagnostics,
                            aes(x = biomass_over_npoc, y = Vh)) +
  geom_point(shape = 1, stroke = 1, size = 2, color = "#b35806") +
  ggpubr::stat_cor(
    aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")),
    label.x.npc = "left",
    label.y.npc = "top",
    vjust = 1
  ) +
  ggpmisc::stat_poly_line(color = "black") +
  ggpubr::theme_pubr(border = TRUE) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 14, face = "bold")
  ) +
  xlab(bquote("Biomass/NPOC")) +
  ylab(bquote("Fitted Vh"))

p_vh_npoc_ssa <- ggplot(df_vh_diagnostics,
                        aes(x = npoc_over_ssa, y = Vh)) +
  geom_point(shape = 1, stroke = 1, size = 2, color = "#b35806") +
  ggpubr::stat_cor(
    aes(label = paste(..rr.label.., ..p.label.., sep = "*`,`~")),
    label.x.npc = "right",
    label.y.npc = "top",
    hjust = 1,
    vjust = 1
  ) +
  ggpmisc::stat_poly_line(color = "black") +
  scale_x_log10(
    name = expression(log[10]~"(NPOC / SSA)"),
    breaks = c(1e-5, 1e-4, 1e-3),
    labels = c(expression(10^-5), expression(10^-4), expression(10^-3))
  ) +
  ggpubr::theme_pubr(border = TRUE) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 14, face = "bold")
  ) +
  ylab(bquote("Fitted Vh"))

write_csv(df_vh_diagnostics, derived_file("fitted_vh_diagnostics.csv"))

ggsave(figure_file("mechanistic_vh_distribution.pdf"), p_vh,
       width = 3.2, height = 3)
ggsave(figure_file("mechanistic_vh_vs_biomass_npoc.pdf"), p_vh_biomass_npoc,
       width = 3.3, height = 3)
ggsave(figure_file("mechanistic_vh_vs_npoc_ssa.pdf"), p_vh_npoc_ssa,
       width = 3.3, height = 3)

if (requireNamespace("patchwork", quietly = TRUE)) {
  p_vh_diagnostics <- p_vh | p_vh_biomass_npoc | p_vh_npoc_ssa
  ggsave(figure_file("mechanistic_revised_figure4_vh_diagnostics.pdf"),
         p_vh_diagnostics, width = 9.8, height = 3)
  p_vh_diagnostics
} else {
  p_vh
  p_vh_biomass_npoc
  p_vh_npoc_ssa
}
```
