---
title: "03 KGML Residual Learning"
author: "Jianqiu Zheng"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
source("R/helpers.R")
library(rsample)
library(ranger)
library(mgcv)
library(caret)
library(xgboost)
```

## Purpose

This notebook tests KGML models that learn the residual between observed respiration and mechanistic predictions.
The mechanistic baseline uses a fixed `Vh0 = 0.10` so that KGML learns
discrepancy from a prescribed accessibility assumption rather than from
sample-specific accessibility values.

## Workflow map

This script is organized into five parts:

1. Build the fixed-`Vh0` mechanistic baseline and residual target.
2. Fit the main KGML models and pure data-driven benchmark models under site-grouped validation.
3. Run diagnostic and sensitivity checks, including NPOC transformations, `Vh0` sensitivity, and Bulk+FTICR upper-bound models.
4. Combine predictions and statistics into reusable CSV tables.
5. Generate manuscript and supplement figures for Section 3.3.

## 1. Fixed-`Vh0` KGML Analysis Table

This chunk reads the mechanistic predictions from script 02 and rebuilds the mechanistic baseline with a prescribed universal `Vh0`.
The residual target for KGML is `dR = observed log10(R) - mechanistic log10(R)`.
Change `fixed_vh0_kgml` here if you want to rerun script 03 with a different fixed accessibility baseline.

```{r load-data}
mech <- read_csv(derived_file("mechanistic_predictions.csv"),
                 col_types = cols(site = col_character()))

fixed_vh0_kgml <- 0.10

kgml_base <- mech %>%
  mutate(
    Vh0_fixed = fixed_vh0_kgml,
    thermo_accessibility = abs(stoichMet_donor_mean) / (Vh0_fixed * npoc),
    R_mech_fixed_vh0 = exp(-thermo_accessibility) * biomass_log10,
    logR_mech_fixed_vh0 = safe_log10(R_mech_fixed_vh0),
    dR_fixed_vh0 = logR_obs - logR_mech_fixed_vh0
  ) %>%
  transmute(
    site, Position,
    respC, npoc, inv_npoc, log_npoc = log10(npoc),
    biomass = biomass_log10,
    delGcat = delGcat_mean,
    lambda = lambda_mean,
    ne = ne_mean,
    stoichMet_donor = stoichMet_donor_mean,
    clay, silt, fine, med, coarse, ssa_tot, d50_mm,
    Vh0_fixed, thermo_accessibility,
    logR_obs,
    logR_mech = logR_mech_fixed_vh0,
    dR = dR_fixed_vh0
  ) %>%
  filter(if_all(c(respC, npoc, inv_npoc, log_npoc, logR_obs, logR_mech, dR),
                is.finite))

nrow(kgml_base)
```

## 2. Cross-Validation and Model-Fitting Helpers

These helper functions are used throughout script 03.
They enforce site-safe validation, perform fold-wise imputation/scaling, and fit RF, XGBoost, and GAM models.
The important design choice is that preprocessing happens inside each training fold to avoid leakage.

```{r grouped-cv-helper}
set.seed(42)

make_group_folds <- function(dat, v = 5, repeats = 10, seed = 42) {
  set.seed(seed)
  group_vfold_cv(dat, group = site, v = v, repeats = repeats)
}

make_loso_folds <- function(dat) {
  sites <- sort(unique(dat$site))
  splits <- lapply(sites, function(s) {
    make_splits(
      list(analysis = which(dat$site != s), assessment = which(dat$site == s)),
      dat
    )
  })

  tibble(splits = splits, id = paste0("LOSO_", sites))
}

fold_id <- function(folds, i) {
  id <- if ("id" %in% names(folds)) as.character(folds$id[[i]]) else paste0("Fold", i)
  if ("id2" %in% names(folds)) id <- paste(id, as.character(folds$id2[[i]]), sep = "_")
  id
}

mean_impute_from_train_03 <- function(train_df, test_df, predictors) {
  train_out <- train_df
  test_out <- test_df

  for (nm in predictors) {
    fill <- mean(train_out[[nm]], na.rm = TRUE)
    if (!is.finite(fill)) fill <- 0
    train_out[[nm]][!is.finite(train_out[[nm]]) | is.na(train_out[[nm]])] <- fill
    test_out[[nm]][!is.finite(test_out[[nm]]) | is.na(test_out[[nm]])] <- fill
  }

  list(train = train_out, test = test_out)
}

scale_from_train_03 <- function(train_df, test_df, predictors) {
  train_out <- train_df
  test_out <- test_df

  for (nm in predictors) {
    center <- mean(train_out[[nm]], na.rm = TRUE)
    spread <- sd(train_out[[nm]], na.rm = TRUE)
    if (!is.finite(center)) center <- 0
    if (!is.finite(spread) || spread == 0) spread <- 1
    train_out[[nm]] <- (train_out[[nm]] - center) / spread
    test_out[[nm]] <- (test_out[[nm]] - center) / spread
  }

  list(train = train_out, test = test_out)
}

fit_rf_cv <- function(dat, predictors, label, folds, response = "dR",
                      prediction_type = c("kgml", "direct")) {
  prediction_type <- match.arg(prediction_type)
  formula <- reformulate(predictors, response = "dR")

  results <- vector("list", nrow(folds))
  importance <- vector("list", nrow(folds))

  for (i in seq_len(nrow(folds))) {
    train_data <- analysis(folds$splits[[i]])
    test_data <- assessment(folds$splits[[i]])

    fixed <- mean_impute_from_train_03(train_data, test_data, predictors)
    formula <- reformulate(predictors, response = response)

    rf_fit <- ranger(
      formula = formula,
      data = fixed$train,
      num.trees = 800,
      min.node.size = 5,
      importance = "permutation",
      seed = 7
    )

    y_hat <- predict(rf_fit, data = fixed$test)$predictions
    logR_pred <- if (prediction_type == "kgml") test_data$logR_mech + y_hat else y_hat

    results[[i]] <- test_data %>%
      mutate(
        fold_id = fold_id(folds, i),
        model = label,
        dR_hat = logR_pred - logR_mech,
        logR_pred = logR_pred
      )

    importance[[i]] <- rf_fit$variable.importance
  }

  list(
    predictions = bind_rows(results),
    importance = importance
  )
}

fit_lm_cv <- function(dat, predictors, label, folds) {
  results <- vector("list", nrow(folds))

  for (i in seq_len(nrow(folds))) {
    train_data <- analysis(folds$splits[[i]])
    test_data <- assessment(folds$splits[[i]])
    fixed <- mean_impute_from_train_03(train_data, test_data, predictors)

    fit <- lm(reformulate(predictors, response = "logR_obs"), data = fixed$train)
    logR_pred <- as.numeric(predict(fit, newdata = fixed$test))

    results[[i]] <- test_data %>%
      mutate(
        fold_id = fold_id(folds, i),
        model = label,
        dR_hat = logR_pred - logR_mech,
        logR_pred = logR_pred
      )
  }

  bind_rows(results)
}

fit_xgb_cv <- function(dat, predictors, label, folds, response = "logR_obs",
                       prediction_type = c("direct", "kgml")) {
  prediction_type <- match.arg(prediction_type)
  results <- vector("list", nrow(folds))

  for (i in seq_len(nrow(folds))) {
    train_data <- analysis(folds$splits[[i]])
    test_data <- assessment(folds$splits[[i]])
    fixed <- mean_impute_from_train_03(train_data, test_data, predictors)
    scaled <- scale_from_train_03(fixed$train, fixed$test, predictors)

    dtrain <- xgb.DMatrix(
      data = as.matrix(scaled$train[, predictors, drop = FALSE]),
      label = scaled$train[[response]]
    )
    dtest <- xgb.DMatrix(data = as.matrix(scaled$test[, predictors, drop = FALSE]))

    set.seed(7)
    fit <- xgb.train(
      params = list(
        objective = "reg:squarederror",
        eta = 0.04,
        max_depth = 2,
        subsample = 0.8,
        colsample_bytree = 0.8,
        min_child_weight = 3
      ),
      data = dtrain,
      nrounds = 120,
      verbose = 0
    )

    y_hat <- as.numeric(predict(fit, dtest))
    logR_pred <- if (prediction_type == "kgml") test_data$logR_mech + y_hat else y_hat

    results[[i]] <- test_data %>%
      mutate(
        fold_id = fold_id(folds, i),
        model = label,
        dR_hat = logR_pred - logR_mech,
        logR_pred = logR_pred
      )
  }

  bind_rows(results)
}

fit_gam_cv <- function(dat, folds) {
  results <- vector("list", nrow(folds))

  for (i in seq_len(nrow(folds))) {
    train_data <- analysis(folds$splits[[i]])
    test_data <- assessment(folds$splits[[i]])

    gam_fit <- gam(dR ~ s(log_npoc, k = 5), data = train_data, method = "REML")
    dR_hat <- predict(gam_fit, newdata = test_data)

    results[[i]] <- test_data %>%
      mutate(
        fold_id = fold_id(folds, i),
        model = "KGML GAM NPOC",
        dR_hat = dR_hat,
        logR_pred = logR_mech + dR_hat
      )
  }

  bind_rows(results)
}
```

## 3. Main KGML Models

The following chunks fit the core KGML models used in Section 3.3.
All KGML models learn `dR`, then add the learned correction back to the fixed-`Vh0` mechanistic prediction.

### 3.1 KGML-NPOC

This is the minimal residual-learning model.
It tests whether the dominant mechanistic bias can be corrected using only NPOC and transformed NPOC terms.

```{r npoc-only-kgml}
folds_grouped <- make_group_folds(kgml_base)
folds_loso <- make_loso_folds(kgml_base)

npoc_predictors <- c("npoc", "inv_npoc", "log_npoc")
rf_npoc <- fit_rf_cv(kgml_base, npoc_predictors, "KGML RF NPOC", folds_grouped)
rf_npoc_loso <- fit_rf_cv(kgml_base, npoc_predictors, "KGML RF NPOC", folds_loso)
pure_rf_npoc <- fit_rf_cv(
  kgml_base, npoc_predictors, "Pure RF NPOC", folds_grouped,
  response = "logR_obs", prediction_type = "direct"
)
pure_xgb_npoc <- fit_xgb_cv(
  kgml_base, npoc_predictors, "Pure XGBoost NPOC", folds_grouped
)
pure_rf_npoc_loso <- fit_rf_cv(
  kgml_base, npoc_predictors, "Pure RF NPOC", folds_loso,
  response = "logR_obs", prediction_type = "direct"
)
pure_xgb_npoc_loso <- fit_xgb_cv(
  kgml_base, npoc_predictors, "Pure XGBoost NPOC", folds_loso
)

cv_npoc <- rf_npoc$predictions
cv_npoc_loso <- rf_npoc_loso$predictions
```

### 3.2 KGML-Bulk

This model expands the residual-learning predictors to include NPOC, biomass, thermodynamic summaries, and grain-size/SSA variables.
It is the main accessibility-focused KGML model.

```{r expanded-kgml}
expanded_predictors <- c(
  "npoc", "inv_npoc", "log_npoc",
  "biomass", "delGcat", "lambda", "ne", "stoichMet_donor",
  "clay", "silt", "fine", "med", "coarse", "ssa_tot", "d50_mm"
)

kgml_expanded <- kgml_base %>%
  filter(if_all(all_of(expanded_predictors), is.finite))

folds_grouped_expanded <- make_group_folds(kgml_expanded)
folds_loso_expanded <- make_loso_folds(kgml_expanded)

rf_expanded <- fit_rf_cv(
  kgml_expanded, expanded_predictors, "KGML RF expanded", folds_grouped_expanded
)
rf_expanded_loso <- fit_rf_cv(
  kgml_expanded, expanded_predictors, "KGML RF expanded", folds_loso_expanded
)
cv_expanded <- rf_expanded$predictions
cv_expanded_loso <- rf_expanded_loso$predictions
```

## 4. Pure Data-Driven Benchmarks

These models predict `log10(R)` directly from the same bulk predictors used by KGML-Bulk.
They are included to evaluate whether KGML is competitive with flexible empirical models and whether residual structure reflects system behavior rather than only the mechanistic model form.

```{r pure-data-driven-models}
pure_rf <- fit_rf_cv(
  kgml_expanded, expanded_predictors, "Pure RF", folds_grouped_expanded,
  response = "logR_obs", prediction_type = "direct"
)
pure_xgb <- fit_xgb_cv(
  kgml_expanded, expanded_predictors, "Pure XGBoost", folds_grouped_expanded
)

pure_rf_loso <- fit_rf_cv(
  kgml_expanded, expanded_predictors, "Pure RF", folds_loso_expanded,
  response = "logR_obs", prediction_type = "direct"
)
pure_xgb_loso <- fit_xgb_cv(
  kgml_expanded, expanded_predictors, "Pure XGBoost", folds_loso_expanded
)
```

## 5. KGML-FTICR and Pure FTICR Benchmarks

This section adds molecular summary features from FTICR-MS.
`KGML-FTICR` learns residuals from NPOC plus FTICR summaries, while `Pure RF-FTICR` and `Pure XGBoost-FTICR` predict `log10(R)` directly from the same FTICR-oriented predictor set.
Correlated FTICR predictors are removed within each training fold.

```{r fticr-kgml}
fticr_path <- raw_file("fticr_features.csv")

if (file.exists(fticr_path)) {
  ft <- read_csv(fticr_path, col_types = cols(site = col_character()))

  fticr_cols <- c(
    "n_peaks", "shannon",
    "NOSC_med", "NOSC_iqr", "NOSC_mean", "NOSC_p10", "NOSC_p90", "NOSC_range90",
    "DBE_med", "DBE_mean", "DBE_iqr",
    "AI_med", "AI_iqr",
    "delG_med", "delG_iqr",
    "lambda_med", "lambda_iqr",
    "Cnum_med", "Cnum_iqr",
    "frac_NOSC_pos", "frac_NOSC_low", "frac_AI_high", "DBE_C_ratio_mean",
    "frac_ConHC", "frac_Lipid", "frac_AminoSugar", "frac_Protein", "frac_Tannin",
    "frac_UnsatHC", "frac_Lignin", "frac_Carb", "frac_Other",
    "ratio_Protein_Lignin", "ratio_Lipid_Carb", "ratio_UnsatHC_ConHC"
  )

  fticr_available <- intersect(fticr_cols, names(ft))

  kgml_ft <- kgml_base %>%
    left_join(ft %>% select(site, Position, all_of(fticr_available)),
              by = c("site", "Position")) %>%
    filter(if_all(all_of(c(npoc_predictors, fticr_available)), is.finite))

  # Correlation filtering is intentionally inside each fold to avoid leakage.
  folds <- make_group_folds(kgml_ft)
  folds_ft_loso <- make_loso_folds(kgml_ft)
  results <- vector("list", nrow(folds))
  importance <- vector("list", nrow(folds))

  for (i in seq_len(nrow(folds))) {
    train_data <- analysis(folds$splits[[i]])
    test_data <- assessment(folds$splits[[i]])

    corr_mat <- cor(train_data[, fticr_available], use = "pairwise.complete.obs")
    high_corr <- findCorrelation(corr_mat, cutoff = 0.80)
    fticr_reduced <- if (length(high_corr) > 0) fticr_available[-high_corr] else fticr_available
    predictors <- c(npoc_predictors, fticr_reduced)
    formula <- reformulate(predictors, response = "dR")

    rf_fit <- ranger(
      formula = formula,
      data = train_data,
      num.trees = 800,
      min.node.size = 5,
      importance = "permutation",
      seed = 7
    )

    dR_hat <- predict(rf_fit, data = test_data)$predictions
    results[[i]] <- test_data %>%
      mutate(
        fold_id = fold_id(folds, i),
        model = "KGML RF FTICR",
        dR_hat = dR_hat,
        logR_pred = logR_mech + dR_hat
      )
    importance[[i]] <- rf_fit$variable.importance
  }

  cv_ft <- bind_rows(results)

  loso_results <- vector("list", nrow(folds_ft_loso))

  for (i in seq_len(nrow(folds_ft_loso))) {
    train_data <- analysis(folds_ft_loso$splits[[i]])
    test_data <- assessment(folds_ft_loso$splits[[i]])

    corr_mat <- cor(train_data[, fticr_available], use = "pairwise.complete.obs")
    high_corr <- findCorrelation(corr_mat, cutoff = 0.80)
    fticr_reduced <- if (length(high_corr) > 0) fticr_available[-high_corr] else fticr_available
    predictors <- c(npoc_predictors, fticr_reduced)
    formula <- reformulate(predictors, response = "dR")

    fixed <- mean_impute_from_train_03(train_data, test_data, predictors)

    rf_fit <- ranger(
      formula = formula,
      data = fixed$train,
      num.trees = 800,
      min.node.size = 5,
      importance = "permutation",
      seed = 7
    )

    dR_hat <- predict(rf_fit, data = fixed$test)$predictions
    loso_results[[i]] <- test_data %>%
      mutate(
        fold_id = fold_id(folds_ft_loso, i),
        model = "KGML RF FTICR",
        dR_hat = dR_hat,
        logR_pred = logR_mech + dR_hat
      )
  }

  cv_ft_loso <- bind_rows(loso_results)

  fit_pure_fticr_cv <- function(dat, label, folds, engine = c("rf", "xgb")) {
    engine <- match.arg(engine)
    results <- vector("list", nrow(folds))

    for (i in seq_len(nrow(folds))) {
      train_data <- analysis(folds$splits[[i]])
      test_data <- assessment(folds$splits[[i]])

      corr_mat <- cor(train_data[, fticr_available], use = "pairwise.complete.obs")
      high_corr <- findCorrelation(corr_mat, cutoff = 0.80)
      fticr_reduced <- if (length(high_corr) > 0) fticr_available[-high_corr] else fticr_available
      predictors <- c(npoc_predictors, fticr_reduced)
      fixed <- mean_impute_from_train_03(train_data, test_data, predictors)

      if (engine == "rf") {
        rf_fit <- ranger(
          formula = reformulate(predictors, response = "logR_obs"),
          data = fixed$train,
          num.trees = 800,
          min.node.size = 5,
          importance = "permutation",
          seed = 7
        )
        logR_pred <- predict(rf_fit, data = fixed$test)$predictions
      } else {
        scaled <- scale_from_train_03(fixed$train, fixed$test, predictors)
        dtrain <- xgb.DMatrix(
          data = as.matrix(scaled$train[, predictors, drop = FALSE]),
          label = scaled$train$logR_obs
        )
        dtest <- xgb.DMatrix(
          data = as.matrix(scaled$test[, predictors, drop = FALSE])
        )

        set.seed(7)
        xgb_fit <- xgb.train(
          params = list(
            objective = "reg:squarederror",
            eta = 0.04,
            max_depth = 2,
            subsample = 0.8,
            colsample_bytree = 0.8,
            min_child_weight = 3
          ),
          data = dtrain,
          nrounds = 120,
          verbose = 0
        )
        logR_pred <- as.numeric(predict(xgb_fit, dtest))
      }

      results[[i]] <- test_data %>%
        mutate(
          fold_id = fold_id(folds, i),
          model = label,
          dR_hat = logR_pred - logR_mech,
          logR_pred = logR_pred
        )
    }

    bind_rows(results)
  }

  pure_rf_fticr <- fit_pure_fticr_cv(
    kgml_ft, "Pure RF-FTICR", folds, engine = "rf"
  )
  pure_xgb_fticr <- fit_pure_fticr_cv(
    kgml_ft, "Pure XGBoost-FTICR", folds, engine = "xgb"
  )
  pure_rf_fticr_loso <- fit_pure_fticr_cv(
    kgml_ft, "Pure RF-FTICR", folds_ft_loso, engine = "rf"
  )
  pure_xgb_fticr_loso <- fit_pure_fticr_cv(
    kgml_ft, "Pure XGBoost-FTICR", folds_ft_loso, engine = "xgb"
  )

  write_csv(cv_ft, derived_file("kgml_fticr_cv_predictions.csv"))
  write_csv(cv_ft_loso, derived_file("kgml_fticr_loso_predictions.csv"))
  write_csv(pure_rf_fticr, derived_file("pure_rf_fticr_cv_predictions.csv"))
  write_csv(pure_xgb_fticr, derived_file("pure_xgboost_fticr_cv_predictions.csv"))
  write_csv(pure_rf_fticr_loso, derived_file("pure_rf_fticr_loso_predictions.csv"))
  write_csv(pure_xgb_fticr_loso, derived_file("pure_xgboost_fticr_loso_predictions.csv"))

  cv_ft
}
```

## 6. Bulk+FTICR Upper-Bound Sensitivity

This sensitivity combines the bulk predictor set and FTICR molecular summaries.
It asks whether KGML improves simply by increasing predictor dimensionality.
Use this section as an upper-bound or supplement analysis, not as the primary model comparison.

```{r bulk-fticr-sensitivity}
if (exists("kgml_ft")) {
  kgml_bulk_ft <- kgml_expanded %>%
    left_join(ft %>% select(site, Position, all_of(fticr_available)),
              by = c("site", "Position")) %>%
    filter(if_all(all_of(fticr_available), is.finite))

  folds_grouped_bulk_ft <- make_group_folds(kgml_bulk_ft)
  folds_loso_bulk_ft <- make_loso_folds(kgml_bulk_ft)

  fit_bulk_fticr_cv <- function(dat, label, folds,
                                response = c("dR", "logR_obs"),
                                prediction_type = c("kgml", "direct"),
                                engine = c("rf", "xgb")) {
    response <- match.arg(response)
    prediction_type <- match.arg(prediction_type)
    engine <- match.arg(engine)
    results <- vector("list", nrow(folds))

    for (i in seq_len(nrow(folds))) {
      train_data <- analysis(folds$splits[[i]])
      test_data <- assessment(folds$splits[[i]])

      corr_mat <- cor(train_data[, fticr_available], use = "pairwise.complete.obs")
      high_corr <- findCorrelation(corr_mat, cutoff = 0.80)
      fticr_reduced <- if (length(high_corr) > 0) {
        fticr_available[-high_corr]
      } else {
        fticr_available
      }
      predictors <- unique(c(expanded_predictors, fticr_reduced))
      fixed <- mean_impute_from_train_03(train_data, test_data, predictors)

      if (engine == "rf") {
        rf_fit <- ranger(
          formula = reformulate(predictors, response = response),
          data = fixed$train,
          num.trees = 800,
          min.node.size = 5,
          importance = "permutation",
          seed = 7
        )
        y_hat <- predict(rf_fit, data = fixed$test)$predictions
      } else {
        scaled <- scale_from_train_03(fixed$train, fixed$test, predictors)
        dtrain <- xgb.DMatrix(
          data = as.matrix(scaled$train[, predictors, drop = FALSE]),
          label = scaled$train[[response]]
        )
        dtest <- xgb.DMatrix(
          data = as.matrix(scaled$test[, predictors, drop = FALSE])
        )

        set.seed(7)
        xgb_fit <- xgb.train(
          params = list(
            objective = "reg:squarederror",
            eta = 0.04,
            max_depth = 2,
            subsample = 0.8,
            colsample_bytree = 0.8,
            min_child_weight = 3
          ),
          data = dtrain,
          nrounds = 120,
          verbose = 0
        )
        y_hat <- as.numeric(predict(xgb_fit, dtest))
      }

      logR_pred <- if (prediction_type == "kgml") test_data$logR_mech + y_hat else y_hat

      results[[i]] <- test_data %>%
        mutate(
          fold_id = fold_id(folds, i),
          model = label,
          dR_hat = logR_pred - logR_mech,
          logR_pred = logR_pred
        )
    }

    bind_rows(results)
  }

  bulk_fticr_predictions <- bind_rows(
    fit_bulk_fticr_cv(
      kgml_bulk_ft, "KGML-Bulk+FTICR", folds_grouped_bulk_ft,
      response = "dR", prediction_type = "kgml", engine = "rf"
    ) %>%
      mutate(Validation = "Site-grouped 5-fold CV"),
    fit_bulk_fticr_cv(
      kgml_bulk_ft, "Pure RF-Bulk+FTICR", folds_grouped_bulk_ft,
      response = "logR_obs", prediction_type = "direct", engine = "rf"
    ) %>%
      mutate(Validation = "Site-grouped 5-fold CV"),
    fit_bulk_fticr_cv(
      kgml_bulk_ft, "Pure XGBoost-Bulk+FTICR", folds_grouped_bulk_ft,
      response = "logR_obs", prediction_type = "direct", engine = "xgb"
    ) %>%
      mutate(Validation = "Site-grouped 5-fold CV"),
    fit_bulk_fticr_cv(
      kgml_bulk_ft, "KGML-Bulk+FTICR", folds_loso_bulk_ft,
      response = "dR", prediction_type = "kgml", engine = "rf"
    ) %>%
      mutate(Validation = "Leave-one-site-out CV"),
    fit_bulk_fticr_cv(
      kgml_bulk_ft, "Pure RF-Bulk+FTICR", folds_loso_bulk_ft,
      response = "logR_obs", prediction_type = "direct", engine = "rf"
    ) %>%
      mutate(Validation = "Leave-one-site-out CV"),
    fit_bulk_fticr_cv(
      kgml_bulk_ft, "Pure XGBoost-Bulk+FTICR", folds_loso_bulk_ft,
      response = "logR_obs", prediction_type = "direct", engine = "xgb"
    ) %>%
      mutate(Validation = "Leave-one-site-out CV")
  ) %>%
    select(Validation, model, fold_id, site, Position, respC, npoc, inv_npoc,
           log_npoc, logR_obs, logR_mech, dR, dR_hat, logR_pred)

  bulk_fticr_stats <- bulk_fticr_predictions %>%
    group_by(Validation, model) %>%
    group_modify(~ model_stats(.x$logR_obs, .x$logR_pred)) %>%
    ungroup() %>%
    rename(Model = model) %>%
    mutate(
      Dataset = "Bulk+FTICR",
      Model_class = case_when(
        str_detect(Model, "^KGML") ~ "KGML",
        TRUE ~ "Pure data-driven"
      )
    ) %>%
    select(Validation, Dataset, Model_class, Model,
           n, RMSE, Pearson_r, CCC, Bias, R2)

  write_csv(bulk_fticr_predictions,
            derived_file("kgml_bulk_fticr_sensitivity_predictions.csv"))
  write_csv(bulk_fticr_stats,
            derived_file("kgml_bulk_fticr_sensitivity_stats.csv"))

  p_bulk_fticr_sensitivity <- bulk_fticr_stats %>%
    select(Validation, Model_class, Model, RMSE, Pearson_r, CCC) %>%
    pivot_longer(c(RMSE, Pearson_r, CCC),
                 names_to = "Metric", values_to = "Value") %>%
    mutate(
      Metric = recode(Metric, Pearson_r = "Pearson r"),
      Metric = factor(Metric, levels = c("RMSE", "Pearson r", "CCC")),
      Model = factor(Model, levels = rev(c(
        "KGML-Bulk+FTICR",
        "Pure RF-Bulk+FTICR",
        "Pure XGBoost-Bulk+FTICR"
      )))
    ) %>%
    ggplot(aes(x = Value, y = Model, color = Model_class)) +
    geom_segment(aes(x = 0, xend = Value, yend = Model),
                 linewidth = 0.6, alpha = 0.75) +
    geom_point(size = 2.4) +
    facet_grid(Validation ~ Metric, scales = "free_x") +
    scale_color_manual(values = c(
      "KGML" = puor$purple,
      "Pure data-driven" = puor$orange
    )) +
    labs(x = NULL, y = NULL, color = NULL) +
    theme_review() +
    theme(
      legend.position = "top",
      panel.grid.major.y = element_blank(),
      strip.text = element_text(face = "bold")
    )

  ggsave(figure_file("kgml_bulk_fticr_sensitivity_lollipop.pdf"),
         p_bulk_fticr_sensitivity, width = 8.5, height = 4.2)

  bulk_fticr_stats
}
```

### 6.1 Bulk+FTICR RF vs GAM Correction Function

This figure compares the RF residual correction against a simpler GAM approximation in the Bulk+FTICR setting.
Because Bulk+FTICR is multidimensional, the curve is projected along `log10(NPOC)` rather than being a true one-variable function.
The GAM uses key bulk predictors plus fold-wise FTICR PCA axes.

```{r bulk-fticr-rf-gam-correction-function}
if (exists("kgml_bulk_ft")) {
  fit_bulk_fticr_gam_correction_cv <- function(dat, folds) {
    gam_bulk_predictors <- c("log_npoc", "biomass", "stoichMet_donor",
                             "ssa_tot", "d50_mm")
    results <- vector("list", nrow(folds))

    for (i in seq_len(nrow(folds))) {
      train_data <- analysis(folds$splits[[i]])
      test_data <- assessment(folds$splits[[i]])

      fixed_fticr <- mean_impute_from_train_03(train_data, test_data,
                                               fticr_available)
      fticr_scaled <- scale_from_train_03(fixed_fticr$train,
                                          fixed_fticr$test,
                                          fticr_available)

      pca_fit <- prcomp(fticr_scaled$train[, fticr_available, drop = FALSE],
                        center = FALSE, scale. = FALSE)
      n_pc <- min(2, ncol(pca_fit$x))
      train_pc <- as.data.frame(pca_fit$x[, seq_len(n_pc), drop = FALSE])
      test_pc <- as.data.frame(
        predict(pca_fit,
                newdata = fticr_scaled$test[, fticr_available, drop = FALSE])[
                  , seq_len(n_pc), drop = FALSE
                ]
      )
      names(train_pc) <- paste0("fticr_pc", seq_len(n_pc))
      names(test_pc) <- paste0("fticr_pc", seq_len(n_pc))

      train_model <- bind_cols(train_data, train_pc)
      test_model <- bind_cols(test_data, test_pc)

      gam_predictors <- c(gam_bulk_predictors, names(train_pc))
      fixed_model <- mean_impute_from_train_03(train_model, test_model,
                                               gam_predictors)

      gam_formula <- reformulate(
        paste0("s(", gam_predictors, ", k = 4)"),
        response = "dR"
      )

      gam_fit <- gam(gam_formula, data = fixed_model$train,
                     method = "REML", select = TRUE)
      dR_hat <- as.numeric(predict(gam_fit, newdata = fixed_model$test))

      results[[i]] <- test_data %>%
        mutate(
          fold_id = fold_id(folds, i),
          model = "GAM Bulk+FTICR",
          dR_hat = dR_hat,
          logR_pred = logR_mech + dR_hat
        )
    }

    bind_rows(results)
  }

  gam_bulk_fticr_correction <- fit_bulk_fticr_gam_correction_cv(
    kgml_bulk_ft, folds_grouped_bulk_ft
  )

  rf_bulk_fticr_correction <- bulk_fticr_predictions %>%
    filter(Validation == "Site-grouped 5-fold CV",
           model == "KGML-Bulk+FTICR")

  bulk_fticr_correction_lines <- bind_rows(
    rf_bulk_fticr_correction %>%
      transmute(site, Position, log_npoc, Model = "RF", dR_hat),
    gam_bulk_fticr_correction %>%
      transmute(site, Position, log_npoc, Model = "GAM", dR_hat)
  ) %>%
    group_by(Model, site, Position) %>%
    summarise(
      log_npoc = mean(log_npoc, na.rm = TRUE),
      dR_hat = mean(dR_hat, na.rm = TRUE),
      .groups = "drop"
    )

  bulk_fticr_correction_smooth <- bulk_fticr_correction_lines %>%
    group_by(Model) %>%
    group_modify(~ {
      x_grid <- tibble(
        log_npoc = seq(min(.x$log_npoc, na.rm = TRUE),
                       max(.x$log_npoc, na.rm = TRUE),
                       length.out = 120)
      )
      fit <- loess(dR_hat ~ log_npoc, span = 0.65, data = .x)
      x_grid %>%
        mutate(dR_hat = as.numeric(predict(fit, newdata = x_grid)))
    }) %>%
    ungroup()

  bulk_fticr_correction_points <- kgml_bulk_ft %>%
    distinct(site, Position, log_npoc, dR)

  write_csv(bulk_fticr_correction_lines,
            derived_file("kgml_bulk_fticr_rf_gam_correction_points.csv"))
  write_csv(bulk_fticr_correction_smooth,
            derived_file("kgml_bulk_fticr_rf_gam_correction_function.csv"))
  write_csv(gam_bulk_fticr_correction,
            derived_file("kgml_bulk_fticr_gam_correction_predictions.csv"))

  p_bulk_fticr_correction <- ggplot() +
    geom_point(
      data = bulk_fticr_correction_points,
      aes(x = log_npoc, y = dR),
      color = "grey65", alpha = 0.55, size = 1.5
    ) +
    geom_line(
      data = bulk_fticr_correction_smooth,
      aes(x = log_npoc, y = dR_hat, color = Model),
      linewidth = 0.9
    ) +
    scale_color_manual(values = c(GAM = puor$purple, RF = puor$orange)) +
    labs(x = expression(log[10](NPOC)),
         y = "Residual correction (dR)",
         color = "Model") +
    theme_review()

  ggsave(figure_file("kgml_bulk_fticr_rf_gam_correction_function.pdf"),
         p_bulk_fticr_correction, width = 4.2, height = 3.4)

  p_bulk_fticr_correction
}
```

## 7. Predictor Manifest

This chunk writes a plain CSV that records each model's dataset, target, sample/site count, and predictor list.
Use this file when checking whether a model is a bulk, FTICR, or Bulk+FTICR comparison.




```{r predictor-manifest}
predictor_manifest <- bind_rows(
  tibble(
    Model = "KGML-NPOC",
    Dataset = "kgml_base",
    Target = "Residual dR",
    samples = nrow(kgml_base),
    sites = n_distinct(kgml_base$site),
    predictors = paste(npoc_predictors, collapse = "; "),
    note = ""
  ),
  tibble(
    Model = "KGML-Bulk",
    Dataset = "kgml_expanded",
    Target = "Residual dR",
    samples = nrow(kgml_expanded),
    sites = n_distinct(kgml_expanded$site),
    predictors = paste(expanded_predictors, collapse = "; "),
    note = ""
  ),
  tibble(
    Model = "Pure RF-Bulk",
    Dataset = "kgml_expanded",
    Target = "Observed log10(R)",
    samples = nrow(kgml_expanded),
    sites = n_distinct(kgml_expanded$site),
    predictors = paste(expanded_predictors, collapse = "; "),
    note = ""
  ),
  tibble(
    Model = "Pure XGBoost-Bulk",
    Dataset = "kgml_expanded",
    Target = "Observed log10(R)",
    samples = nrow(kgml_expanded),
    sites = n_distinct(kgml_expanded$site),
    predictors = paste(expanded_predictors, collapse = "; "),
    note = ""
  )
)

if (exists("kgml_ft")) {
  predictor_manifest <- bind_rows(
    predictor_manifest,
    tibble(
      Model = "KGML-FTICR",
      Dataset = "kgml_ft",
      Target = "Residual dR",
      samples = nrow(kgml_ft),
      sites = n_distinct(kgml_ft$site),
      predictors = paste(c(npoc_predictors, fticr_available), collapse = "; "),
      note = "FTICR predictors are filtered for high correlation within each training fold."
    ),
    tibble(
      Model = "Pure RF-FTICR",
      Dataset = "kgml_ft",
      Target = "Observed log10(R)",
      samples = nrow(kgml_ft),
      sites = n_distinct(kgml_ft$site),
      predictors = paste(c(npoc_predictors, fticr_available), collapse = "; "),
      note = "FTICR predictors are filtered for high correlation within each training fold."
    ),
    tibble(
      Model = "Pure XGBoost-FTICR",
      Dataset = "kgml_ft",
      Target = "Observed log10(R)",
      samples = nrow(kgml_ft),
      sites = n_distinct(kgml_ft$site),
      predictors = paste(c(npoc_predictors, fticr_available), collapse = "; "),
      note = "FTICR predictors are filtered for high correlation within each training fold."
    )
  )
}

write_csv(predictor_manifest,
          derived_file("kgml_model_predictor_manifest.csv"))

predictor_manifest
```

## 8. NPOC Term Sensitivity

This diagnostic tests how much KGML-Bulk and KGML-FTICR depend on NPOC-related predictors.
It compares three predictor settings: no NPOC terms, raw `npoc` only, and the full transformed NPOC predictor set (`npoc`, `inv_npoc`, and `log_npoc`).
The output figure is useful as a supplement for showing whether model performance is driven by the NPOC terms alone.

```{r npoc-transform-ablation}
npoc_transform_ablation_predictions <- bind_rows(
  fit_rf_cv(
    kgml_expanded,
    expanded_predictors,
    "KGML-Bulk full NPOC transforms",
    folds_grouped_expanded
  )$predictions %>%
    mutate(Validation = "Site-grouped 5-fold CV"),
  fit_rf_cv(
    kgml_expanded,
    setdiff(expanded_predictors, c("inv_npoc", "log_npoc")),
    "KGML-Bulk raw NPOC only",
    folds_grouped_expanded
  )$predictions %>%
    mutate(Validation = "Site-grouped 5-fold CV"),
  fit_rf_cv(
    kgml_expanded,
    setdiff(expanded_predictors, c("npoc", "inv_npoc", "log_npoc")),
    "KGML-Bulk no NPOC terms",
    folds_grouped_expanded
  )$predictions %>%
    mutate(Validation = "Site-grouped 5-fold CV"),
  fit_rf_cv(
    kgml_expanded,
    expanded_predictors,
    "KGML-Bulk full NPOC transforms",
    folds_loso_expanded
  )$predictions %>%
    mutate(Validation = "Leave-one-site-out CV"),
  fit_rf_cv(
    kgml_expanded,
    setdiff(expanded_predictors, c("inv_npoc", "log_npoc")),
    "KGML-Bulk raw NPOC only",
    folds_loso_expanded
  )$predictions %>%
    mutate(Validation = "Leave-one-site-out CV"),
  fit_rf_cv(
    kgml_expanded,
    setdiff(expanded_predictors, c("npoc", "inv_npoc", "log_npoc")),
    "KGML-Bulk no NPOC terms",
    folds_loso_expanded
  )$predictions %>%
    mutate(Validation = "Leave-one-site-out CV")
)

if (exists("kgml_ft")) {
  fticr_full_predictors <- c(npoc_predictors, fticr_available)
  fticr_raw_npoc_predictors <- c("npoc", fticr_available)
  fticr_no_npoc_predictors <- fticr_available

  fit_fticr_ablation <- function(dat, predictors, label, folds) {
    results <- vector("list", nrow(folds))

    for (i in seq_len(nrow(folds))) {
      train_data <- analysis(folds$splits[[i]])
      test_data <- assessment(folds$splits[[i]])

      fticr_in_model <- intersect(fticr_available, predictors)
      if (length(fticr_in_model) > 1) {
        corr_mat <- cor(train_data[, fticr_in_model], use = "pairwise.complete.obs")
        high_corr <- findCorrelation(corr_mat, cutoff = 0.80)
        fticr_reduced <- if (length(high_corr) > 0) {
          fticr_in_model[-high_corr]
        } else {
          fticr_in_model
        }
      } else {
        fticr_in_model
      }

      predictors_fold <- unique(c(intersect(c("npoc", "inv_npoc", "log_npoc"),
                                            predictors),
                                  fticr_reduced))
      fixed <- mean_impute_from_train_03(train_data, test_data, predictors_fold)

      rf_fit <- ranger(
        formula = reformulate(predictors_fold, response = "dR"),
        data = fixed$train,
        num.trees = 800,
        min.node.size = 5,
        importance = "permutation",
        seed = 7
      )

      dR_hat <- predict(rf_fit, data = fixed$test)$predictions
      results[[i]] <- test_data %>%
        mutate(
          fold_id = fold_id(folds, i),
          model = label,
          dR_hat = dR_hat,
          logR_pred = logR_mech + dR_hat
        )
    }

    bind_rows(results)
  }

  npoc_transform_ablation_predictions <- bind_rows(
    npoc_transform_ablation_predictions,
    fit_fticr_ablation(
      kgml_ft, fticr_full_predictors,
      "KGML-FTICR full NPOC transforms", folds
    ) %>%
      mutate(Validation = "Site-grouped 5-fold CV"),
    fit_fticr_ablation(
      kgml_ft, fticr_raw_npoc_predictors,
      "KGML-FTICR raw NPOC only", folds
    ) %>%
      mutate(Validation = "Site-grouped 5-fold CV"),
    fit_fticr_ablation(
      kgml_ft, fticr_no_npoc_predictors,
      "KGML-FTICR no NPOC terms", folds
    ) %>%
      mutate(Validation = "Site-grouped 5-fold CV"),
    fit_fticr_ablation(
      kgml_ft, fticr_full_predictors,
      "KGML-FTICR full NPOC transforms", folds_ft_loso
    ) %>%
      mutate(Validation = "Leave-one-site-out CV"),
    fit_fticr_ablation(
      kgml_ft, fticr_raw_npoc_predictors,
      "KGML-FTICR raw NPOC only", folds_ft_loso
    ) %>%
      mutate(Validation = "Leave-one-site-out CV"),
    fit_fticr_ablation(
      kgml_ft, fticr_no_npoc_predictors,
      "KGML-FTICR no NPOC terms", folds_ft_loso
    ) %>%
      mutate(Validation = "Leave-one-site-out CV")
  )
}

npoc_transform_ablation_predictions <- npoc_transform_ablation_predictions %>%
  select(Validation, model, fold_id, site, Position, respC, npoc, inv_npoc,
         log_npoc, logR_obs, logR_mech, dR, dR_hat, logR_pred)

npoc_transform_ablation_stats <- npoc_transform_ablation_predictions %>%
  group_by(Validation, model) %>%
  group_modify(~ model_stats(.x$logR_obs, .x$logR_pred)) %>%
  ungroup() %>%
  rename(Model = model) %>%
  mutate(
    Model_family = case_when(
      str_detect(Model, "Bulk") ~ "KGML-Bulk",
      str_detect(Model, "FTICR") ~ "KGML-FTICR",
      TRUE ~ Model
    ),
    NPOC_setting = case_when(
      str_detect(Model, "no NPOC") ~ "no NPOC terms",
      str_detect(Model, "full") ~ "npoc + inv_npoc + log_npoc",
      str_detect(Model, "raw") ~ "npoc only",
      TRUE ~ Model
    )
  ) %>%
  select(Validation, Model_family, NPOC_setting, Model,
         n, RMSE, Pearson_r, CCC, Bias, R2)

write_csv(npoc_transform_ablation_predictions,
          derived_file("kgml_npoc_transform_ablation_predictions.csv"))
write_csv(npoc_transform_ablation_stats,
          derived_file("kgml_npoc_transform_ablation_stats.csv"))
write_csv(npoc_transform_ablation_stats,
          derived_file("kgml_npoc_term_sensitivity_stats.csv"))

p_npoc_transform_ablation <- npoc_transform_ablation_stats %>%
  select(Validation, Model_family, NPOC_setting, RMSE, Pearson_r, CCC) %>%
  pivot_longer(c(RMSE, Pearson_r, CCC),
               names_to = "Metric", values_to = "Value") %>%
  mutate(
    Metric = recode(Metric, Pearson_r = "Pearson r"),
    Metric = factor(Metric, levels = c("RMSE", "Pearson r", "CCC")),
    NPOC_setting = factor(NPOC_setting,
                          levels = c("no NPOC terms", "npoc only",
                                     "npoc + inv_npoc + log_npoc"))
  ) %>%
  ggplot(aes(x = NPOC_setting, y = Value, color = Model_family,
             group = Model_family)) +
  geom_line(linewidth = 0.65) +
  geom_point(size = 2.2) +
  facet_grid(Validation ~ Metric, scales = "free_y") +
  scale_color_manual(values = c(
    "KGML-Bulk" = puor$orange,
    "KGML-FTICR" = puor$purple
  )) +
  labs(x = NULL, y = NULL, color = NULL) +
  theme_review() +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

ggsave(figure_file("kgml_npoc_transform_ablation.pdf"),
       p_npoc_transform_ablation, width = 8.5, height = 5.2)
ggsave(figure_file("kgml_npoc_term_sensitivity.pdf"),
       p_npoc_transform_ablation, width = 8.5, height = 5.2)

npoc_transform_ablation_stats
```


## 9. GAM-NPOC Comparison

This chunk fits a simple GAM residual model using `log_npoc`.
It is used mainly for the supplement figure comparing RF and GAM correction shapes for the NPOC-only discrepancy.

```{r gam-comparison}
cv_gam <- fit_gam_cv(kgml_base, folds_grouped)
cv_gam_loso <- fit_gam_cv(kgml_base, folds_loso)
```

## 10. Fixed-`Vh0` Sensitivity for NPOC-Only KGML

This sensitivity reruns NPOC-only residual learning across several prescribed fixed `Vh0` values.
It helps separate the effect of the chosen universal accessibility baseline from the learned residual correction.

```{r npoc-vh0-sensitivity}
vh0_sensitivity_values <- c(
  fixed_0.05 = 0.05,
  fixed_0.10 = 0.10,
  fixed_0.25 = 0.25,
  fixed_0.50 = 0.50,
  fixed_1.00 = 1.00
)

make_kgml_base_for_vh0 <- function(vh0) {
  mech %>%
    mutate(
      Vh0_fixed = vh0,
      thermo_accessibility = abs(stoichMet_donor_mean) / (Vh0_fixed * npoc),
      R_mech_fixed_vh0 = exp(-thermo_accessibility) * biomass_log10,
      logR_mech_fixed_vh0 = safe_log10(R_mech_fixed_vh0),
      dR_fixed_vh0 = logR_obs - logR_mech_fixed_vh0
    ) %>%
    transmute(
      site, Position,
      respC, npoc, inv_npoc, log_npoc = log10(npoc),
      biomass = biomass_log10,
      delGcat = delGcat_mean,
      lambda = lambda_mean,
      ne = ne_mean,
      stoichMet_donor = stoichMet_donor_mean,
      clay, silt, fine, med, coarse, ssa_tot, d50_mm,
      Vh0_fixed, thermo_accessibility,
      logR_obs,
      logR_mech = logR_mech_fixed_vh0,
      dR = dR_fixed_vh0
    ) %>%
    filter(if_all(c(respC, npoc, inv_npoc, log_npoc, logR_obs, logR_mech, dR),
                  is.finite))
}

fit_gam_npoc_original_cv <- function(dat, folds) {
  results <- vector("list", nrow(folds))

  for (i in seq_len(nrow(folds))) {
    train_data <- analysis(folds$splits[[i]])
    test_data <- assessment(folds$splits[[i]])

    gam_fit <- gam(
      dR ~ s(npoc, k = 5) + s(inv_npoc, k = 5) + s(log_npoc, k = 5),
      data = train_data,
      method = "REML"
    )

    dR_hat <- predict(gam_fit, newdata = test_data)

    results[[i]] <- test_data %>%
      mutate(
        fold_id = fold_id(folds, i),
        model = "KGML GAM NPOC original",
        dR_hat = dR_hat,
        logR_pred = logR_mech + dR_hat
      )
  }

  bind_rows(results)
}

npoc_vh0_sensitivity_predictions <- purrr::imap_dfr(
  vh0_sensitivity_values,
  function(vh0, vh0_label) {
    dat <- make_kgml_base_for_vh0(vh0)
    grouped <- make_group_folds(dat)
    loso <- make_loso_folds(dat)

    bind_rows(
      fit_rf_cv(dat, npoc_predictors, "KGML-NPOC RF",
                grouped)$predictions %>%
        mutate(Validation = "Site-grouped 5-fold CV"),
      fit_gam_npoc_original_cv(dat, grouped) %>%
        mutate(Validation = "Site-grouped 5-fold CV"),
      fit_rf_cv(dat, npoc_predictors, "KGML-NPOC RF",
                loso)$predictions %>%
        mutate(Validation = "Leave-one-site-out CV"),
      fit_gam_npoc_original_cv(dat, loso) %>%
        mutate(Validation = "Leave-one-site-out CV")
    ) %>%
      mutate(
        Vh0_label = vh0_label,
        Vh0 = vh0,
        dR_hat = as.numeric(dR_hat),
        logR_pred = as.numeric(logR_pred)
      ) %>%
      select(Validation, Vh0_label, Vh0, model, fold_id, site, Position,
             respC, npoc, inv_npoc, log_npoc, logR_obs, logR_mech,
             dR, dR_hat, logR_pred)
  }
)

npoc_vh0_sensitivity_stats <- npoc_vh0_sensitivity_predictions %>%
  group_by(Validation, Vh0_label, Vh0, model) %>%
  group_modify(~ model_stats(.x$logR_obs, .x$logR_pred)) %>%
  ungroup() %>%
  rename(Model = model) %>%
  select(Validation, Vh0_label, Vh0, Model, n, RMSE, Pearson_r, CCC, Bias, R2)

write_csv(npoc_vh0_sensitivity_predictions,
          derived_file("kgml_npoc_vh0_sensitivity_predictions.csv"))
write_csv(npoc_vh0_sensitivity_stats,
          derived_file("kgml_npoc_vh0_sensitivity_stats.csv"))

p_npoc_vh0_sensitivity <- npoc_vh0_sensitivity_stats %>%
  select(Validation, Vh0_label, Vh0, Model, RMSE, Pearson_r, CCC) %>%
  pivot_longer(c(RMSE, Pearson_r, CCC),
               names_to = "Metric", values_to = "Value") %>%
  mutate(
    Metric = recode(Metric, Pearson_r = "Pearson r"),
    Metric = factor(Metric, levels = c("RMSE", "Pearson r", "CCC")),
    Vh0_label = factor(Vh0_label, levels = names(vh0_sensitivity_values))
  ) %>%
  ggplot(aes(x = Vh0_label, y = Value, color = Model, group = Model)) +
  geom_line(linewidth = 0.65) +
  geom_point(size = 2) +
  facet_grid(Validation ~ Metric, scales = "free_y") +
  scale_color_manual(values = c(
    "KGML-NPOC RF" = puor$orange,
    "KGML GAM NPOC original" = puor$purple
  )) +
  labs(x = expression(fixed~V[h][0]), y = NULL, color = NULL) +
  theme_review() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(figure_file("kgml_npoc_vh0_sensitivity.pdf"),
       p_npoc_vh0_sensitivity, width = 8.5, height = 5.2)

npoc_vh0_sensitivity_stats
```

## 11. Combined Prediction and Statistics Tables

This chunk standardizes the prediction outputs from all main KGML and pure data-driven models.
It writes the core CSV files used by the Section 3.3 performance figures:
`kgml_cv_predictions_all.csv`, `kgml_cv_stats_all.csv`, and `kgml_cv_stats_compact.csv`.

```{r combined-kgml-stats}
standardize_predictions <- function(dat, validation, model_label) {
  dat %>%
    mutate(
      Validation = validation,
      Vh0 = fixed_vh0_kgml,
      Model = model_label
    ) %>%
    select(Validation, Vh0, Model, fold_id, site, Position, respC, npoc, inv_npoc,
           log_npoc, logR_obs, logR_mech, dR, dR_hat, logR_pred) %>%
    mutate(across(c(respC, npoc, inv_npoc, log_npoc, logR_obs, logR_mech,
                    dR, dR_hat, logR_pred), as.numeric))
}

kgml_cv_predictions_all <- bind_rows(
  standardize_predictions(cv_npoc, "Site-grouped 5-fold CV", "KGML-NPOC"),
  standardize_predictions(pure_rf_npoc$predictions, "Site-grouped 5-fold CV", "Pure RF-NPOC"),
  standardize_predictions(pure_xgb_npoc, "Site-grouped 5-fold CV", "Pure XGBoost-NPOC"),
  standardize_predictions(cv_expanded, "Site-grouped 5-fold CV", "KGML-Bulk"),
  if (exists("cv_ft")) standardize_predictions(cv_ft, "Site-grouped 5-fold CV", "KGML-FTICR"),
  standardize_predictions(cv_gam, "Site-grouped 5-fold CV", "KGML-GAM NPOC"),
  standardize_predictions(pure_rf$predictions, "Site-grouped 5-fold CV", "Pure RF-Bulk"),
  standardize_predictions(pure_xgb, "Site-grouped 5-fold CV", "Pure XGBoost-Bulk"),
  if (exists("pure_rf_fticr")) standardize_predictions(pure_rf_fticr, "Site-grouped 5-fold CV", "Pure RF-FTICR"),
  if (exists("pure_xgb_fticr")) standardize_predictions(pure_xgb_fticr, "Site-grouped 5-fold CV", "Pure XGBoost-FTICR"),
  standardize_predictions(cv_npoc_loso, "Leave-one-site-out CV", "KGML-NPOC"),
  standardize_predictions(pure_rf_npoc_loso$predictions, "Leave-one-site-out CV", "Pure RF-NPOC"),
  standardize_predictions(pure_xgb_npoc_loso, "Leave-one-site-out CV", "Pure XGBoost-NPOC"),
  standardize_predictions(cv_expanded_loso, "Leave-one-site-out CV", "KGML-Bulk"),
  if (exists("cv_ft_loso")) standardize_predictions(cv_ft_loso, "Leave-one-site-out CV", "KGML-FTICR"),
  standardize_predictions(cv_gam_loso, "Leave-one-site-out CV", "KGML-GAM NPOC"),
  standardize_predictions(pure_rf_loso$predictions, "Leave-one-site-out CV", "Pure RF-Bulk"),
  standardize_predictions(pure_xgb_loso, "Leave-one-site-out CV", "Pure XGBoost-Bulk"),
  if (exists("pure_rf_fticr_loso")) standardize_predictions(pure_rf_fticr_loso, "Leave-one-site-out CV", "Pure RF-FTICR"),
  if (exists("pure_xgb_fticr_loso")) standardize_predictions(pure_xgb_fticr_loso, "Leave-one-site-out CV", "Pure XGBoost-FTICR")
)

mechanistic_predictions_all <- bind_rows(
  cv_npoc %>%
    mutate(Validation = "Site-grouped 5-fold CV", Vh0 = fixed_vh0_kgml,
           Model = "Mechanistic fixed Vh0", logR_pred = logR_mech,
           dR_hat = 0) %>%
    select(Validation, Vh0, Model, fold_id, site, Position, respC, npoc,
           inv_npoc, log_npoc, logR_obs, logR_mech, dR, dR_hat, logR_pred),
  cv_npoc_loso %>%
    mutate(Validation = "Leave-one-site-out CV", Vh0 = fixed_vh0_kgml,
           Model = "Mechanistic fixed Vh0", logR_pred = logR_mech,
           dR_hat = 0) %>%
    select(Validation, Vh0, Model, fold_id, site, Position, respC, npoc,
           inv_npoc, log_npoc, logR_obs, logR_mech, dR, dR_hat, logR_pred)
)

kgml_cv_predictions_all <- bind_rows(
  mechanistic_predictions_all,
  kgml_cv_predictions_all
)

kgml_cv_stats_all <- kgml_cv_predictions_all %>%
  group_by(Validation, Vh0, Model) %>%
  group_modify(~ model_stats(.x$logR_obs, .x$logR_pred)) %>%
  ungroup() %>%
  select(Validation, Vh0, Model, everything()) %>%
  arrange(
    factor(Validation, levels = c("Site-grouped 5-fold CV", "Leave-one-site-out CV")),
    RMSE
  )

write_csv(kgml_cv_predictions_all,
          derived_file("kgml_cv_predictions_all.csv"))
write_csv(kgml_cv_stats_all,
          derived_file("kgml_cv_stats_all.csv"))

kgml_cv_stats_compact <- kgml_cv_stats_all %>%
  select(Validation, Vh0, Model, n, RMSE, Pearson_r, CCC, Bias)

write_csv(kgml_cv_stats_compact,
          derived_file("kgml_cv_stats_compact.csv"))

kgml_cv_stats_all
```


##############################################################################################################
## 12. Figure: KGML Residual Structure

This chunk creates the main residual-structure plots for KGML-NPOC, KGML-Bulk, and KGML-FTICR.
Each plot contrasts the original mechanistic residual with the KGML residual along `1/NPOC`.
```{r combined-figures}
plot_inputs <- list(
  cv_npoc %>% select(site, Position, logR_obs, logR_mech, logR_pred, npoc, model),
  cv_expanded %>% select(site, Position, logR_obs, logR_mech, logR_pred, npoc, model),
  cv_gam %>% select(site, Position, logR_obs, logR_mech, logR_pred, npoc, model)
)

if (exists("cv_ft")) {
  plot_inputs <- c(
    plot_inputs,
    list(cv_ft %>% select(site, Position, logR_obs, logR_mech,
                          logR_pred, npoc, model))
  )
}

plot_data <- bind_rows(plot_inputs) %>%
  group_by(model, site, Position) %>%
  summarise(
    logR_obs = mean(logR_obs, na.rm = TRUE),
    logR_mech = mean(logR_mech, na.rm = TRUE),
    logR_pred = mean(logR_pred, na.rm = TRUE),
    npoc = mean(npoc, na.rm = TRUE),
    .groups = "drop"
  )

lims <- range(c(plot_data$logR_obs, plot_data$logR_mech, plot_data$logR_pred),
              na.rm = TRUE)

p_fit <- ggplot(plot_data, aes(x = logR_obs, y = logR_pred, color = model)) +
  geom_point(alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_equal(xlim = lims, ylim = lims) +
  labs(x = "Observed log10(R)", y = "KGML predicted log10(R)", color = "Model") +
  theme_review()

p_fit

resid_data <- plot_data %>%
  mutate(
    inv_npoc = 1 / npoc,
    resid_mech = logR_obs - logR_mech,
    resid_kgml = logR_obs - logR_pred
  ) %>%
  pivot_longer(c(resid_mech, resid_kgml),
               names_to = "series", values_to = "residual") %>%
  mutate(
    model = recode(
      model,
      "KGML RF NPOC" = "KGML-NPOC",
      "KGML RF expanded" = "KGML-Bulk",
      "KGML RF FTICR" = "KGML-FTICR",
      "KGML GAM NPOC" = "KGML-GAM NPOC"
    ),
    series = recode(
      series,
      resid_mech = "Mechanistic",
      resid_kgml = "KGML"
    )
  )

write_csv(resid_data, derived_file("kgml_residual_structure_plot_data.csv"))

make_residual_plot <- function(dat, model_label) {
  dat %>%
    filter(model == model_label) %>%
    ggplot(aes(x = inv_npoc, y = residual, color = series)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(alpha = 0.55) +
    ylim(-2.2, 4) +
    geom_smooth(method = "loess", se = FALSE) +
    scale_color_manual(values = c(
      "Mechanistic" = puor$purple,
      "KGML" = puor$orange
    )) +
    labs(x = "1/NPOC", y = "Residual log10(R)", color = NULL) +
    theme_review()
}

p_resid_npoc <- make_residual_plot(resid_data, "KGML-NPOC")
p_resid_bulk <- make_residual_plot(resid_data, "KGML-Bulk")
p_resid_fticr <- make_residual_plot(resid_data, "KGML-FTICR")

if (requireNamespace("patchwork", quietly = TRUE)) {
  p_resid <- p_resid_npoc | p_resid_bulk | p_resid_fticr
    p_resid <- p_resid +
    patchwork::plot_layout(guides = "collect") &
    theme(legend.position = "right")
  ggsave(figure_file("kgml_residual_structure_main.pdf"), p_resid,
         width = 10, height = 3)
} else {
  p_resid <- p_resid_npoc
}

```

## 13. Supplement Inputs: RF vs GAM NPOC Correction Function

This chunk builds the NPOC correction-function objects used in the four-panel supplement figure below.
It compares the NPOC-only RF residual correction with a GAM correction.
It tests whether the learned residual correction along `log10(NPOC)` is a smooth, systematic pattern rather than an arbitrary RF artifact.
The second subplot shows how the GAM correction changes residual structure against `1/NPOC`.

```{r rf-gam-npoc-correction-figure}
npoc_correction_points <- kgml_base %>%
  distinct(site, Position, log_npoc, dR)

npoc_correction_lines <- bind_rows(
  cv_npoc %>%
    transmute(site, Position, log_npoc, Model = "RF", dR_hat),
  cv_gam %>%
    transmute(site, Position, log_npoc, Model = "GAM", dR_hat)) %>%
  group_by(Model, site, Position) %>%
  summarise(
    log_npoc = mean(log_npoc, na.rm = TRUE),
    dR_hat = mean(dR_hat, na.rm = TRUE),
    .groups = "drop") %>%
  arrange(Model, log_npoc)

write_csv(npoc_correction_lines,
          derived_file("kgml_rf_gam_npoc_correction_function.csv"))

p_npoc_correction <- ggplot() +
  geom_point(
    data = npoc_correction_points,
    aes(x = log_npoc, y = dR),
    color = "grey65", alpha = 0.65, size = 1.6) +
  geom_line(
    data = npoc_correction_lines,
    aes(x = log_npoc, y = dR_hat, color = Model),
    linewidth = 0.9, alpha=0.75) +
  scale_color_manual(values = c(GAM = puor$dark_purple, RF = puor$dark_orange)) +
  labs(x = expression(log[10](NPOC)),
       y = "Residual correction (dR)",
       color = "Model") +
  theme_review()

gam_residual_data <- cv_gam %>%
  group_by(site, Position) %>%
  summarise(
    logR_obs = mean(logR_obs, na.rm = TRUE),
    logR_mech = mean(logR_mech, na.rm = TRUE),
    logR_pred = mean(logR_pred, na.rm = TRUE),
    npoc = mean(npoc, na.rm = TRUE),
    .groups = "drop") %>%
  mutate(
    inv_npoc = 1 / npoc,
    `Mechanistic residual` = logR_obs - logR_mech,
    `GAM residual` = logR_obs - logR_pred) %>%
  pivot_longer(
    c(`Mechanistic residual`, `GAM residual`),
    names_to = "Series",
    values_to = "Residual")

write_csv(gam_residual_data,
          derived_file("kgml_gam_npoc_residual_plot_data.csv"))

p_gam_residual <- ggplot(gam_residual_data,
                         aes(x = inv_npoc, y = Residual, color = Series)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.55, size = 1.6) +
  ylim(-2.2, 4) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 0.8) +
  scale_color_manual(values = c(
    "Mechanistic residual" = puor$purple,
    "GAM residual" = puor$orange)) +
  labs(x = "1/NPOC", y = "Residual log10(R)", color = NULL) +
  theme_review()


if (requireNamespace("patchwork", quietly = TRUE)) {
  p_rf_gam_supplement <- p_gam_residual|p_npoc_correction
  ggsave(figure_file("kgml_rf_gam_npoc_supplement.pdf"),
         p_rf_gam_supplement, width = 5.8, height = 3.3)
}

p_npoc_correction
```

## 14. Supplement Figure: Bulk+FTICR Residual Correction

This supplement figure mirrors the RF-vs-GAM NPOC diagnostic above, but for the Bulk+FTICR sensitivity model.
The left panel shows whether KGML-Bulk+FTICR reduces the mechanistic residual structure along `1/NPOC`.
The right panel shows the projected RF and GAM residual-correction functions along `log10(NPOC)`.

```{r kgml-residual-correction-supplement}
bulk_fticr_predictions <- read_csv(
  derived_file("kgml_bulk_fticr_sensitivity_predictions.csv"),
  col_types = cols(site = col_character())
)

bulk_fticr_correction_smooth <- read_csv(
  derived_file("kgml_bulk_fticr_rf_gam_correction_function.csv"),
  col_types = cols()
)

bulk_fticr_residual_data <- bulk_fticr_predictions %>%
  filter(Validation == "Site-grouped 5-fold CV",
         model == "KGML-Bulk+FTICR") %>%
  group_by(site, Position) %>%
  summarise(
    logR_obs = mean(logR_obs, na.rm = TRUE),
    logR_mech = mean(logR_mech, na.rm = TRUE),
    logR_pred = mean(logR_pred, na.rm = TRUE),
    npoc = mean(npoc, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    inv_npoc = 1 / npoc,
    Mechanistic = logR_obs - logR_mech,
    `KGML-Bulk+FTICR` = logR_obs - logR_pred
  ) %>%
  pivot_longer(
    c(Mechanistic, `KGML-Bulk+FTICR`),
    names_to = "Series",
    values_to = "Residual"
  )

bulk_fticr_correction_points <- bulk_fticr_predictions %>%
  filter(Validation == "Site-grouped 5-fold CV",
         model == "KGML-Bulk+FTICR") %>%
  distinct(site, Position, log_npoc, dR)

write_csv(bulk_fticr_residual_data,
          derived_file("kgml_bulk_fticr_residual_plot_data.csv"))

p_supp_bulk_fticr_residual <- ggplot(
  bulk_fticr_residual_data,
  aes(x = inv_npoc, y = Residual, color = Series)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.55, size = 1.4) +
  ylim(-2.2, 4) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 0.8) +
  scale_color_manual(values = c(
    "Mechanistic" = puor$purple,
    "KGML-Bulk+FTICR" = puor$orange
  )) +
  labs(x = "1/NPOC", y = "Residual log10(R)", color = NULL) +
  theme_review()

p_supp_bulk_fticr_decay <- ggplot() +
  geom_point(
    data = bulk_fticr_correction_points,
    aes(x = log_npoc, y = dR),
    color = "grey65", alpha = 0.55, size = 1.4
  ) +
  geom_line(
    data = bulk_fticr_correction_smooth,
    aes(x = log_npoc, y = dR_hat, color = Model),
    linewidth = 0.9, alpha = 0.75
  ) +
  scale_color_manual(values = c(GAM = puor$dark_purple, RF = puor$dark_orange)) +
  labs(x = expression(log[10](NPOC)),
       y = "Residual correction (dR)",
       color = "Model") +
  theme_review()

if (requireNamespace("patchwork", quietly = TRUE)) {
  p_residual_correction_supplement <- p_supp_bulk_fticr_residual|p_supp_bulk_fticr_decay

  ggsave(figure_file("kgml_residual_correction_supplement.pdf"),
         p_residual_correction_supplement, width = 5.8, height = 3.3)
} else {
  p_residual_correction_supplement <- p_supp_bulk_fticr_residual
}

p_residual_correction_supplement
```

## 16. Figure: Pure Data-Driven Prediction and Residual Structure

This figure compares KGML-Bulk/FTICR against pure RF/XGBoost models using the site-grouped 5-fold predictions.
The first panel set shows observed-vs-predicted agreement; the second shows residual structure along `1/NPOC`.
Use this figure to support the Section 3.3 comparison between mechanistically constrained KGML and flexible pure data-driven prediction.

```{r pure-data-driven-prediction-figure}
pure_kgml_models <- c(
  "KGML-NPOC",
  "Pure RF-NPOC",
  "Pure XGBoost-NPOC",
  "KGML-Bulk",
  "Pure RF-Bulk",
  "Pure XGBoost-Bulk",
  "KGML-FTICR",
  "Pure RF-FTICR",
  "Pure XGBoost-FTICR"
)

pure_prediction_plot_data <- kgml_cv_predictions_all %>%
  filter(
    Validation == "Site-grouped 5-fold CV",
    Model %in% pure_kgml_models
  ) %>%
  group_by(Model, site, Position) %>%
  summarise(
    logR_obs = mean(logR_obs, na.rm = TRUE),
    logR_mech = mean(logR_mech, na.rm = TRUE),
    logR_pred = mean(logR_pred, na.rm = TRUE),
    inv_npoc = mean(inv_npoc, na.rm = TRUE),
    mechanistic_residual = mean(logR_obs - logR_mech, na.rm = TRUE),
    model_residual = mean(logR_obs - logR_pred, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Model = factor(Model, levels = pure_kgml_models),
    Model_class = case_when(
      str_detect(as.character(Model), "^KGML") ~ "KGML",
      TRUE ~ "Pure data-driven"
    ),
    Predictor_set = case_when(
      str_detect(as.character(Model), "NPOC") ~ "NPOC",
      str_detect(as.character(Model), "Bulk") ~ "Bulk",
      str_detect(as.character(Model), "FTICR") ~ "FTICR",
      TRUE ~ NA_character_
    ),
    Model_type = case_when(
      str_detect(as.character(Model), "^KGML") ~ "KGML",
      str_detect(as.character(Model), "RF") ~ "Pure RF",
      str_detect(as.character(Model), "XGBoost") ~ "Pure XGBoost",
      TRUE ~ as.character(Model)
    )
  )

pure_prediction_plot_data <- pure_prediction_plot_data %>%
  mutate(Predictor_set = factor(Predictor_set, levels = c("NPOC", "Bulk", "FTICR")))

pure_residual_plot_data <- pure_prediction_plot_data %>%
  pivot_longer(
    c(mechanistic_residual, model_residual),
    names_to = "Series",
    values_to = "Residual"
  ) %>%
  mutate(
    Series = recode(
      Series,
      mechanistic_residual = "Mechanistic",
      model_residual = "Model"
    )
  )

write_csv(pure_prediction_plot_data,
          derived_file("kgml_pure_data_prediction_plot_data.csv"))
write_csv(pure_residual_plot_data,
          derived_file("kgml_pure_data_residual_plot_data.csv"))

pure_model_colors <- c(
  "KGML" = puor$purple,
  "Pure RF" = puor$orange,
  "Pure XGBoost" = puor$dark_orange
)

lims_pure <- range(c(pure_prediction_plot_data$logR_obs,
                     pure_prediction_plot_data$logR_pred),
                   na.rm = TRUE)

p_pure_prediction <- ggplot(
  pure_prediction_plot_data,
  aes(x = logR_obs, y = logR_pred, color = Model_type)
) +
  geom_point(alpha = 0.62, size = 1.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_equal(xlim = lims_pure, ylim = lims_pure) +
  facet_wrap(~ Predictor_set, nrow = 1) +
  scale_color_manual(values = pure_model_colors) +
  labs(x = "Observed log10(R)", y = "Predicted log10(R)",
       color = NULL) +
  theme_review() +
  theme(legend.position = "top")

p_pure_residual <- ggplot(
  pure_residual_plot_data,
  aes(x = inv_npoc, y = Residual, color = Series)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.55, size = 1.4) +
  ylim(-2.2, 4) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 0.8) +
  facet_wrap(~ Model, nrow = 2) +
  scale_color_manual(values = c(
    "Mechanistic" = puor$purple,
    "Model" = puor$orange
  )) +
  labs(x = "1/NPOC", y = "Residual log10(R)") +
  theme_review() +
  theme(legend.position = "none")

ggsave(figure_file("kgml_pure_data_prediction_1to1.pdf"),
       p_pure_prediction, width = 7.2, height = 3.6)
ggsave(figure_file("kgml_pure_data_residual_structure.pdf"),
       p_pure_residual, width = 8.6, height = 5.6)

p_pure_prediction
p_pure_residual
```

## 17. Figure: Model Performance Lollipop by Validation Scheme

These figures compare model performance across RMSE, Pearson r, and CCC.
KGML and pure tree-based data-driven models are shown together.
Site-grouped 5-fold and leave-one-site-out results are saved as separate figures rather than faceted by validation scheme.

```{r combined-stats-lollipop}
metric_labels <- c(
  RMSE = "RMSE",
  Pearson_r = "Pearson r",
  CCC = "CCC"
)

lollipop_data <- kgml_cv_stats_all %>%
  filter(Model != "KGML-GAM NPOC") %>%
  select(Validation, Model, RMSE, Pearson_r, CCC) %>%
  pivot_longer(c(RMSE, Pearson_r, CCC),
               names_to = "Metric", values_to = "Value") %>%
  mutate(
    Metric = factor(metric_labels[Metric],
                    levels = c("RMSE", "Pearson r", "CCC")),
    Model = factor(Model, levels = rev(c(
      "Mechanistic fixed Vh0",
      "KGML-NPOC",
      "Pure RF-NPOC",
      "Pure XGBoost-NPOC",
      "KGML-Bulk",
      "Pure RF-Bulk",
      "Pure XGBoost-Bulk",
      "KGML-FTICR",
      "Pure RF-FTICR",
      "Pure XGBoost-FTICR"
    )))
  )

write_csv(lollipop_data,
          derived_file("kgml_model_comparison_lollipop_data.csv"))

make_validation_lollipop <- function(dat, validation_label) {
  dat %>%
    filter(Validation == validation_label) %>%
    ggplot(aes(x = Value, y = Model, color = Model)) +
    geom_segment(aes(x = 0, xend = Value, yend = Model),
                 linewidth = 1, alpha = 0.85) +
    geom_point(size = 2.3) +
    facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
    scale_color_puor(levels(lollipop_data$Model)) +
    labs(x = NULL, y = NULL) +
    theme_review() +
    theme(
      legend.position = "none",
      panel.grid.major.y = element_blank(),
      strip.text = element_text(face = "bold")
    )
}

p_stats_lollipop_grouped <- make_validation_lollipop(
  lollipop_data, "Site-grouped 5-fold CV"
)
p_stats_lollipop_loso <- make_validation_lollipop(
  lollipop_data, "Leave-one-site-out CV"
)

ggsave(figure_file("kgml_tree_model_comparison_grouped5fold.pdf"),
       p_stats_lollipop_grouped, width = 8, height = 3.6)
ggsave(figure_file("kgml_tree_model_comparison_loso.pdf"),
       p_stats_lollipop_loso, width = 8, height = 3.6)

p_stats_lollipop_grouped
p_stats_lollipop_loso
```
