---
title: "04 Clean Validation and Vh0 Sensitivity"
author: "Jianqiu Zheng"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
source("R/helpers.R")
library(rsample)
library(ranger)
library(xgboost)
library(caret)

set.seed(42)
rf_trees <- 600
```

## Purpose

This workflow rebuilds validation and fixed-`Vh0` sensitivity from the curated
logic in scripts 01-03. The rule for the residual learner is strict:

- The mechanistic model contributes through `logR_mech` and the residual target
  `dR = logR_obs - logR_mech`.
- KGML residual predictors are measured variables or direct measurement-derived
  descriptors only.
- Mechanistic intermediate terms and products involving mechanistic predictions
  are not predictors.

The tested accessibility scales are fixed `Vh0` scenarios from `0.05` to
`1.00`.

## 1. Data and Predictor Sets

This section loads the curated mechanistic and FTICR data, defines fixed `Vh0` scenarios, and declares the measured predictor sets used throughout the validation workflow.

```{r load-data}
mech <- read_csv(derived_file("mechanistic_predictions.csv"),
                 col_types = cols(site = col_character()))

ft <- read_csv(raw_file("fticr_features.csv"),
               col_types = cols(site = col_character()))

fticr_cols <- c(
  "n_peaks", "shannon",
  "NOSC_med", "NOSC_iqr", "NOSC_mean", "NOSC_p10", "NOSC_p90",
  "NOSC_range90", "DBE_med", "DBE_mean", "DBE_iqr", "AI_med", "AI_iqr",
  "delG_med", "delG_iqr", "lambda_med", "lambda_iqr", "Cnum_med",
  "Cnum_iqr", "frac_NOSC_pos", "frac_NOSC_low", "frac_AI_high",
  "DBE_C_ratio_mean", "frac_ConHC", "frac_Lipid", "frac_AminoSugar",
  "frac_Protein", "frac_Tannin", "frac_UnsatHC", "frac_Lignin",
  "frac_Carb", "frac_Other", "ratio_Protein_Lignin",
  "ratio_Lipid_Carb", "ratio_UnsatHC_ConHC"
)

fticr_available <- intersect(fticr_cols, names(ft))

analysis_base <- mech %>%
  select(-any_of(fticr_available)) %>%
  left_join(
    ft %>% select(site, Position, all_of(fticr_available)),
    by = c("site", "Position")
  ) %>%
  mutate(
    log_npoc = log10(npoc),
    inv_npoc = 1 / npoc,
    logR_obs = safe_log10(respC)
  ) %>%
  filter(
    complete.cases(site, Position, respC, npoc, log_npoc, inv_npoc,
                   biomass_log10, stoichMet_donor_mean, Vh),
    respC > 0, npoc > 0
  )

vh0_grid <- tibble(
  Vh0_label = c("0.05", "0.10", "0.25", "0.50", "1.00"),
  Vh0 = c(0.05, 0.10, 0.25, 0.50, 1.00),
  Vh0_display = Vh0_label
)

npoc_predictors <- c("npoc", "inv_npoc", "log_npoc")

bulk_predictors <- c(
  "npoc", "inv_npoc", "log_npoc",
  "biomass_log10", "delGcat_mean", "lambda_mean", "ne_mean",
  "stoichMet_donor_mean", "clay", "silt", "fine", "med", "coarse",
  "ssa_tot", "d50_mm"
)

analysis_base %>%
  summarise(
    samples = n(),
    sites = n_distinct(site)
  )
```

## 2. Validation Helpers

This section defines reusable cross-validation, preprocessing, and model-fitting helpers used by the validation and sensitivity analyses.

```{r helpers}
make_group_folds_clean <- function(dat, v = 5, repeats = 10, seed = 42) {
  set.seed(seed)
  group_vfold_cv(dat, group = site, v = v, repeats = repeats)
}

make_loso_folds_clean <- function(dat) {
  sites <- sort(unique(dat$site))
  splits <- lapply(sites, function(s) {
    make_splits(
      list(analysis = which(dat$site != s), assessment = which(dat$site == s)),
      dat
    )
  })

  tibble(splits = splits, id = paste0("LOSO_", sites))
}

fold_id_clean <- function(folds, i) {
  id <- if ("id" %in% names(folds)) as.character(folds$id[[i]]) else paste0("Fold", i)
  if ("id2" %in% names(folds)) id <- paste(id, as.character(folds$id2[[i]]), sep = "_")
  id
}

mean_impute_from_train_clean <- function(train_df, test_df, predictors) {
  train_out <- train_df
  test_out <- test_df

  for (nm in predictors) {
    fill <- mean(train_out[[nm]], na.rm = TRUE)
    if (!is.finite(fill)) fill <- 0
    train_out[[nm]][!is.finite(train_out[[nm]]) | is.na(train_out[[nm]])] <- fill
    test_out[[nm]][!is.finite(test_out[[nm]]) | is.na(test_out[[nm]])] <- fill
  }

  list(train = train_out, test = test_out)
}

scale_from_train_clean <- function(train_df, test_df, predictors) {
  train_out <- train_df
  test_out <- test_df

  for (nm in predictors) {
    center <- mean(train_out[[nm]], na.rm = TRUE)
    spread <- sd(train_out[[nm]], na.rm = TRUE)
    if (!is.finite(center)) center <- 0
    if (!is.finite(spread) || spread == 0) spread <- 1
    train_out[[nm]] <- (train_out[[nm]] - center) / spread
    test_out[[nm]] <- (test_out[[nm]] - center) / spread
  }

  list(train = train_out, test = test_out)
}

prune_fticr_from_train <- function(train_df, candidates, cutoff = 0.80) {
  candidates <- candidates[candidates %in% names(train_df)]
  candidates <- candidates[vapply(train_df[candidates], is.numeric, logical(1))]

  if (length(candidates) < 2) return(candidates)

  complete_counts <- vapply(train_df[candidates], function(x) {
    sum(is.finite(x))
  }, numeric(1))
  candidates <- candidates[complete_counts >= 5]
  if (length(candidates) < 2) return(candidates)

  corr_mat <- suppressWarnings(cor(train_df[, candidates, drop = FALSE],
                                   use = "pairwise.complete.obs"))
  corr_mat[!is.finite(corr_mat)] <- 0
  high_corr <- caret::findCorrelation(corr_mat, cutoff = cutoff)
  if (length(high_corr) > 0) candidates[-high_corr] else candidates
}

add_fixed_vh0_mechanistic <- function(dat, fixed_vh0) {
  dat %>%
    mutate(
      Vh0 = fixed_vh0,
      R_mech = exp(-abs(stoichMet_donor_mean) / (fixed_vh0 * npoc)) *
        biomass_log10,
      logR_mech = safe_log10(R_mech),
      dR = logR_obs - logR_mech
    )
}

fit_rf_predict <- function(train_df, test_df, response, predictors) {
  predictors <- predictors[predictors %in% names(train_df)]
  fixed <- mean_impute_from_train_clean(train_df, test_df, predictors)

  fit <- ranger(
    formula = reformulate(predictors, response = response),
    data = fixed$train,
    num.trees = rf_trees,
    min.node.size = 5,
    importance = "permutation",
    seed = 7
  )

  predict(fit, data = fixed$test)$predictions
}

fit_xgb_predict <- function(train_df, test_df, response, predictors) {
  predictors <- predictors[predictors %in% names(train_df)]
  fixed <- mean_impute_from_train_clean(train_df, test_df, predictors)
  scaled <- scale_from_train_clean(fixed$train, fixed$test, predictors)

  dtrain <- xgb.DMatrix(
    data = as.matrix(scaled$train[, predictors, drop = FALSE]),
    label = scaled$train[[response]]
  )
  dtest <- xgb.DMatrix(
    data = as.matrix(scaled$test[, predictors, drop = FALSE])
  )

  set.seed(7)
  fit <- xgb.train(
    params = list(
      objective = "reg:squarederror",
      eta = 0.04,
      max_depth = 2,
      subsample = 0.8,
      colsample_bytree = 0.8,
      min_child_weight = 3
    ),
    data = dtrain,
    nrounds = 120,
    verbose = 0
  )

  as.numeric(predict(fit, dtest))
}

predictor_spec <- tribble(
  ~Predictor_set, ~base_predictors, ~include_fticr,
  "NPOC", list(npoc_predictors), FALSE,
  "Bulk", list(bulk_predictors), FALSE,
  "FTICR", list(npoc_predictors), TRUE,
  "Bulk+FTICR", list(bulk_predictors), TRUE
)

predictor_set_levels <- c("NPOC", "Bulk", "FTICR", "Bulk+FTICR")
predictor_set_colors <- c(
  "NPOC" = puor$light_purple,
  "Bulk" = puor$purple,
  "FTICR" = puor$orange,
  "Bulk+FTICR" = puor$dark_orange
)
```

## 3. Model Runners

This section wraps the pure-ML and KGML model fits so they can be run consistently across predictor sets and validation folds.

```{r model-runners}
complete_for_predictor_set <- function(dat, base_predictors, include_fticr) {
  required <- base_predictors
  if (include_fticr) required <- c(required, fticr_available)

  dat %>%
    filter(if_all(all_of(required), is.finite))
}

fold_predictors <- function(train_df, base_predictors, include_fticr) {
  if (!include_fticr) return(base_predictors)
  unique(c(base_predictors, prune_fticr_from_train(train_df, fticr_available)))
}

run_pure_models <- function(dat, folds, validation_label, predictor_set,
                            base_predictors, include_fticr) {
  out <- vector("list", nrow(folds))

  for (i in seq_len(nrow(folds))) {
    train_df <- analysis(folds$splits[[i]])
    test_df <- assessment(folds$splits[[i]])
    predictors <- fold_predictors(train_df, base_predictors, include_fticr)
    fold_id <- fold_id_clean(folds, i)

    rf_pred <- fit_rf_predict(train_df, test_df, "logR_obs", predictors)
    xgb_pred <- fit_xgb_predict(train_df, test_df, "logR_obs", predictors)

    out[[i]] <- bind_rows(
      test_df %>%
        transmute(
          site, Position, fold_id = fold_id,
          Validation = validation_label,
          Predictor_set = predictor_set,
          Model = paste("Pure RF", predictor_set),
          Vh0_label = NA_character_,
          Vh0 = NA_real_,
          logR_obs,
          logR_pred = rf_pred
        ),
      test_df %>%
        transmute(
          site, Position, fold_id = fold_id,
          Validation = validation_label,
          Predictor_set = predictor_set,
          Model = paste("Pure XGBoost", predictor_set),
          Vh0_label = NA_character_,
          Vh0 = NA_real_,
          logR_obs,
          logR_pred = xgb_pred
        )
    )
  }

  bind_rows(out)
}

run_kgml_models <- function(dat, folds, validation_label, predictor_set,
                            base_predictors, include_fticr,
                            fixed_vh0, vh0_label) {
  dat_vh <- add_fixed_vh0_mechanistic(dat, fixed_vh0)
  out <- vector("list", nrow(folds))

  for (i in seq_len(nrow(folds))) {
    train_df <- analysis(folds$splits[[i]]) %>%
      add_fixed_vh0_mechanistic(fixed_vh0)
    test_df <- assessment(folds$splits[[i]]) %>%
      add_fixed_vh0_mechanistic(fixed_vh0)
    predictors <- fold_predictors(train_df, base_predictors, include_fticr)
    fold_id <- fold_id_clean(folds, i)

    dR_hat <- fit_rf_predict(train_df, test_df, "dR", predictors)

    out[[i]] <- bind_rows(
      test_df %>%
        transmute(
          site, Position, fold_id = fold_id,
          Validation = validation_label,
          Predictor_set = predictor_set,
          Model = "Mechanistic",
          Vh0_label = vh0_label,
          Vh0 = fixed_vh0,
          logR_obs,
          logR_pred = logR_mech
        ),
      test_df %>%
        transmute(
          site, Position, fold_id = fold_id,
          Validation = validation_label,
          Predictor_set = predictor_set,
          Model = paste("KGML RF", predictor_set),
          Vh0_label = vh0_label,
          Vh0 = fixed_vh0,
          logR_obs,
          logR_pred = logR_mech + dR_hat
        )
    )
  }

  bind_rows(out)
}
```

## 4. Run Validation and Sensitivity

This section executes grouped and leave-one-site-out validation for each predictor set and writes the combined predictions and statistics.

```{r run-validation}
run_for_predictor_set <- function(spec_row) {
  predictor_set <- spec_row$Predictor_set
  base_predictors <- unlist(spec_row$base_predictors[[1]], use.names = FALSE)
  include_fticr <- spec_row$include_fticr

  dat <- complete_for_predictor_set(
    analysis_base, base_predictors, include_fticr
  )

  folds_grouped <- make_group_folds_clean(dat, v = 5, repeats = 10, seed = 42)
  folds_loso <- make_loso_folds_clean(dat)

  pure_preds <- bind_rows(
    run_pure_models(dat, folds_grouped, "Grouped 5-fold CV",
                    predictor_set, base_predictors, include_fticr),
    run_pure_models(dat, folds_loso, "Leave-one-site-out CV",
                    predictor_set, base_predictors, include_fticr)
  )

  kgml_preds <- pmap_dfr(vh0_grid, function(Vh0_label, Vh0, Vh0_display) {
    bind_rows(
      run_kgml_models(dat, folds_grouped, "Grouped 5-fold CV",
                      predictor_set, base_predictors, include_fticr,
                      fixed_vh0 = Vh0, vh0_label = Vh0_label),
      run_kgml_models(dat, folds_loso, "Leave-one-site-out CV",
                      predictor_set, base_predictors, include_fticr,
                      fixed_vh0 = Vh0, vh0_label = Vh0_label)
    )
  })

  bind_rows(pure_preds, kgml_preds)
}

clean_validation_predictions <- map_dfr(
  seq_len(nrow(predictor_spec)),
  ~ run_for_predictor_set(predictor_spec[.x, ])
)

write_csv(clean_validation_predictions,
          derived_file("clean_validation_sensitivity_predictions.csv"))

clean_validation_stats <- clean_validation_predictions %>%
  group_by(Validation, Predictor_set, Model, Vh0_label, Vh0) %>%
  group_modify(~ model_stats(.x$logR_obs, .x$logR_pred)) %>%
  ungroup() %>%
  arrange(Validation, Predictor_set, Model, Vh0)

write_csv(clean_validation_stats,
          derived_file("clean_validation_sensitivity_stats.csv"))

clean_validation_stats
```

## 5. KGML Versus Pure ML Competition

This section compares each KGML fixed-`Vh0` model against the best pure-ML model within the same predictor set and validation design.

```{r competition}
pure_best <- clean_validation_stats %>%
  filter(str_detect(Model, "^Pure ")) %>%
  group_by(Validation, Predictor_set) %>%
  slice_min(RMSE, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(
    Validation,
    Predictor_set,
    best_pure_model = Model,
    best_pure_RMSE = RMSE
  )

kgml_competition <- clean_validation_stats %>%
  filter(str_detect(Model, "^KGML ")) %>%
  left_join(pure_best, by = c("Validation", "Predictor_set")) %>%
  mutate(
    delta_RMSE_vs_best_pure = RMSE - best_pure_RMSE,
    beats_best_pure = delta_RMSE_vs_best_pure < 0,
    Vh0_label = factor(Vh0_label, levels = vh0_grid$Vh0_label)
  ) %>%
  arrange(Validation, Predictor_set, Vh0)

write_csv(kgml_competition,
          derived_file("clean_validation_kgml_vs_pure_competition.csv"))

vh0_axis_labels <- setNames(vh0_grid$Vh0_display, vh0_grid$Vh0_label)

kgml_competition <- kgml_competition %>%
  mutate(
    Predictor_set = factor(Predictor_set, levels = predictor_set_levels)
  )

p_clean_competition <- kgml_competition %>%
  mutate(
    Line_type = if_else(Predictor_set == "Bulk+FTICR",
                        "Bulk+FTICR", "Primary predictor sets")
  ) %>%
  ggplot(aes(x = Vh0_label, y = delta_RMSE_vs_best_pure,
             color = Predictor_set, group = Predictor_set,
             linetype = Line_type)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.6) +
  facet_wrap(~ Validation) +
  scale_x_discrete(labels = vh0_axis_labels) +
  scale_color_manual(
    values = predictor_set_colors,
    breaks = predictor_set_levels
  ) +
  scale_linetype_manual(
    values = c("Primary predictor sets" = "solid", "Bulk+FTICR" = "dashed"),
    guide = "none"
  ) +
labs(
  x = "Fixed Vh0",
  y = expression(Delta * "RMSE vs best pure ML"),
  color = "Predictor set"
) +
  theme_review() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
    legend.position = "right"
  )

ggsave(figure_file("clean_validation_vh0_kgml_vs_pure_all_predictor_sets.pdf"),
       p_clean_competition, width = 7.8, height = 3.0)

p_clean_competition_main <- kgml_competition %>%
  filter(Predictor_set %in% c("NPOC", "Bulk", "FTICR")) %>%
  mutate(
    Predictor_set = factor(Predictor_set, levels = c("NPOC", "Bulk", "FTICR"))
  ) %>%
  ggplot(aes(x = Vh0_label, y = delta_RMSE_vs_best_pure,
             color = Predictor_set, group = Predictor_set)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.6) +
  facet_wrap(~ Validation) +
  scale_x_discrete(labels = vh0_axis_labels) +
  scale_color_manual(
    values = predictor_set_colors[c("NPOC", "Bulk", "FTICR")],
    breaks = c("NPOC", "Bulk", "FTICR")
  ) +
labs(
  x = "Fixed Vh0",
  y = expression(Delta * "RMSE vs best pure ML"),
  color = "Predictor set"
) +
  theme_review() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
    legend.position = "right"
  )

ggsave(figure_file("clean_validation_vh0_kgml_vs_pure.pdf"),
       p_clean_competition_main, width = 7.8, height = 3.0)
ggsave(figure_file("clean_validation_vh0_kgml_vs_pure_main_text.pdf"),
       p_clean_competition_main, width = 7.8, height = 3.0)

p_clean_competition
p_clean_competition_main
kgml_competition %>% filter(beats_best_pure)
```

## 6. RMSE Sensitivity Summary

This section summarizes how fixed `Vh0` changes KGML and mechanistic RMSE across validation designs.

```{r rmse-sensitivity}
plot_stats <- clean_validation_stats %>%
  filter(
    str_detect(Model, "^KGML ") |
      (Model == "Mechanistic" & Predictor_set == "Bulk")
  ) %>%
  mutate(
    Vh0_label = factor(Vh0_label, levels = vh0_grid$Vh0_label),
    Model_short = case_when(
      Model == "Mechanistic" ~ "Mechanistic",
      TRUE ~ paste("KGML", Predictor_set)
    )
  )

p_clean_rmse <- ggplot(
  plot_stats,
  aes(x = Vh0_label, y = RMSE, color = Model_short, group = Model_short)
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~ Validation) +
  scale_x_discrete(labels = vh0_axis_labels) +
  scale_color_manual(values = c(
    "Mechanistic" = puor$pale_purple,
    "KGML NPOC" = puor$light_purple,
    "KGML Bulk" = puor$purple,
    "KGML FTICR" = puor$orange,
    "KGML Bulk+FTICR" = puor$dark_orange
  )) +
  labs(x = "Fixed Vh0", y = "Validation RMSE", color = NULL) +
  theme_review() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
    legend.position = "bottom"
  )

ggsave(figure_file("clean_validation_vh0_rmse_sensitivity.pdf"),
       p_clean_rmse, width = 8.4, height = 4.8)

p_clean_rmse
```

## 7. Multi-Metric Lollipop Summary

This section reshapes the selected validation metrics into a compact lollipop figure comparing KGML and pure-ML performance.

```{r metric-lollipop-summary}
kgml_metric_summary <- clean_validation_stats %>%
  filter(str_detect(Model, "^KGML "), Vh0 %in% c(0.50, 1.00)) %>%
  mutate(
    Model_group = "KGML RF",
    Selected_model = paste0(Model, " (Vh0 = ", Vh0_label, ")")
  )

best_pure_by_rmse <- clean_validation_stats %>%
  filter(str_detect(Model, "^Pure ")) %>%
  group_by(Validation, Predictor_set) %>%
  slice_min(RMSE, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    Model_group = "Best pure ML",
    Selected_model = Model
  )

metric_lollipop_summary <- bind_rows(
  kgml_metric_summary,
  best_pure_by_rmse
) %>%
  mutate(
    Model_label = case_when(
      Model_group == "Mechanistic" ~ Selected_model,
      TRUE ~ paste(Predictor_set, Selected_model, sep = ": ")
    ),
    Model_group = factor(
      Model_group,
      levels = c("KGML RF", "Best pure ML")
    ),
    Predictor_set = factor(
      Predictor_set,
      levels = predictor_set_levels
    )
  ) %>%
  select(
    Validation, Predictor_set, Model_group, Model_label, Selected_model,
    Vh0_label, Vh0, n, RMSE, Pearson_r, CCC, Bias
  )

write_csv(metric_lollipop_summary,
          derived_file("clean_validation_metric_lollipop_summary.csv"))

metric_lollipop_long <- metric_lollipop_summary %>%
  pivot_longer(
    cols = c(RMSE, Pearson_r, CCC),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  mutate(
    Metric = factor(
      Metric,
      levels = c("RMSE", "Pearson_r", "CCC"),
      labels = c("RMSE", "Pearson r", "CCC")
    ),
    Model_label = factor(Model_label, levels = rev(unique(Model_label)))
  )

write_csv(metric_lollipop_long,
          derived_file("clean_validation_metric_lollipop_long.csv"))

p_metric_lollipop <- metric_lollipop_long %>%
  ggplot(aes(x = Value, y = Model_label, color = Model_group)) +
  geom_vline(xintercept = 0, color = "grey80", linewidth = 0.3) +
  geom_segment(aes(x = 0, xend = Value, yend = Model_label),
               linewidth = 0.65, alpha = 0.75) +
  geom_point(size = 2.2) +
  facet_grid(Metric ~ Validation, scales = "free_x") +
  scale_color_manual(values = c(
    "KGML RF" = puor$purple,
    "Best pure ML" = puor$orange
  )) +
  labs(x = NULL, y = NULL, color = NULL) +
  theme_review() +
  theme(
    strip.text.y = element_text(angle = 0),
    axis.text.y = element_text(size = 7),
    legend.position = "bottom"
  )

ggsave(figure_file("clean_validation_metric_lollipop.pdf"),
       p_metric_lollipop, width = 9.2, height = 8.2)

p_metric_lollipop
metric_lollipop_summary
```

## 8. Accessibility-Scale Importance Comparison

This section estimates permutation importance for KGML residual models and aggregates features into interpretable predictor families.

```{r section34-importance-comparison}
feature_labels <- c(
  npoc = "NPOC",
  inv_npoc = "1/NPOC",
  log_npoc = "log10(NPOC)",
  biomass_log10 = "log10(biomass)",
  delGcat_mean = "catabolic energy",
  lambda_mean = "lambda",
  ne_mean = "electron transfer",
  stoichMet_donor_mean = "donor stoichiometry",
  d50_mm = "d50",
  n_peaks = "peak count",
  frac_Other = "other compound fraction",
  ratio_Lipid_Carb = "lipid:carbohydrate",
  lambda_iqr = "lambda IQR",
  delG_med = "median delG"
)

section34_importance_comparison_grid <- vh0_grid %>%
  filter(Vh0 %in% c(0.25, 0.50, 1.00))

section34_importance_comparison <- tidyr::crossing(
  section34_importance_comparison_grid,
  predictor_spec %>% filter(Predictor_set %in% c("Bulk", "FTICR", "Bulk+FTICR"))
) %>%
  pmap_dfr(function(Vh0_label, Vh0, Vh0_display,
                    Predictor_set, base_predictors, include_fticr) {
    base_predictors <- unlist(base_predictors, use.names = FALSE)
    dat <- complete_for_predictor_set(
      analysis_base, base_predictors, include_fticr
    ) %>%
      add_fixed_vh0_mechanistic(Vh0)

    predictors <- fold_predictors(dat, base_predictors, include_fticr)
    fixed <- mean_impute_from_train_clean(dat, dat, predictors)$train

    fit <- ranger(
      formula = reformulate(predictors, response = "dR"),
      data = fixed,
      num.trees = rf_trees,
      min.node.size = 5,
      importance = "permutation",
      seed = 7
    )

    tibble(
      Predictor_set = Predictor_set,
      Vh0_label = Vh0_label,
      Vh0 = Vh0,
      Vh0_display = Vh0_display,
      Feature = names(fit$variable.importance),
      Importance = as.numeric(fit$variable.importance)
    )
  }) %>%
  group_by(Predictor_set, Vh0_label, Vh0) %>%
  mutate(
    rank = rank(-Importance, ties.method = "first"),
    Feature_label = recode(Feature, !!!feature_labels, .default = Feature)
  ) %>%
  ungroup()

section34_importance_comparison_features <- section34_importance_comparison %>%
  group_by(Predictor_set, Feature, Feature_label) %>%
  summarise(
    mean_rank = mean(rank, na.rm = TRUE),
    max_importance = max(Importance, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(Predictor_set) %>%
  slice_min(mean_rank, n = 8, with_ties = FALSE) %>%
  ungroup()

section34_importance_comparison_top <- section34_importance_comparison %>%
  semi_join(
    section34_importance_comparison_features,
    by = c("Predictor_set", "Feature", "Feature_label")
  ) %>%
  left_join(
    section34_importance_comparison_features,
    by = c("Predictor_set", "Feature", "Feature_label")
  ) %>%
  arrange(Predictor_set, desc(max_importance), Vh0) %>%
  mutate(
    Vh0_label = factor(Vh0_label,
                       levels = section34_importance_comparison_grid$Vh0_label),
    Vh0_display = factor(Vh0_display,
                         levels = section34_importance_comparison_grid$Vh0_display),
    Feature_plot = factor(
      paste(Predictor_set, Feature_label, sep = ": "),
      levels = unique(paste(Predictor_set, Feature_label, sep = ": "))
    )
  )

write_csv(section34_importance_comparison,
          derived_file("section34_vh0_025_050_100_kgml_importance.csv"))
write_csv(section34_importance_comparison_top,
          derived_file("section34_vh0_025_050_100_kgml_importance_top.csv"))

classify_feature_family <- function(feature) {
  case_when(
    feature %in% c("npoc", "inv_npoc", "log_npoc") ~ "NPOC",
    feature %in% c(
      "delGcat_mean", "lambda_mean", "ne_mean", "stoichMet_donor_mean",
      "delG_med", "delG_iqr", "lambda_med", "lambda_iqr",
      "NOSC_med", "NOSC_iqr", "NOSC_mean", "NOSC_p10", "NOSC_p90",
      "NOSC_range90"
    ) ~ "Thermodynamic",
    feature %in% c(
      "clay", "silt", "fine", "med", "coarse", "ssa_tot", "d50_mm"
    ) ~ "Sediment physical structure",
    feature %in% c("biomass_log10") ~ "Biomass",
    str_detect(
      feature,
      "^(n_peaks|shannon|DBE|AI|Cnum|frac_|ratio_|DBE_C_ratio)"
    ) ~ "OM molecular composition",
    TRUE ~ "Other"
  )
}

feature_family_levels <- c(
  "NPOC", "Thermodynamic", "Sediment physical structure",
  "Biomass", "OM molecular composition"
)

feature_family_colors <- c(
  "NPOC" = puor$dark_purple,
  "Thermodynamic" = puor$purple,
  "Sediment physical structure" = puor$light_purple,
  "Biomass" = puor$pale_purple,
  "OM molecular composition" = puor$orange
)

section34_importance_family <- section34_importance_comparison %>%
  mutate(
    Feature_family = classify_feature_family(Feature),
    Importance_positive = pmax(Importance, 0)
  ) %>%
  filter(Feature_family != "Other") %>%
  group_by(Predictor_set, Vh0_label, Vh0, Vh0_display, Feature_family) %>%
  summarise(
    family_importance = sum(Importance_positive, na.rm = TRUE),
    n_features = dplyr::n(),
    .groups = "drop"
  ) %>%
  group_by(Predictor_set, Vh0_label, Vh0, Vh0_display) %>%
  mutate(
    family_share = family_importance / sum(family_importance, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    Vh0_display = sprintf("%.2f", Vh0),
    Vh0_display = factor(Vh0_display, levels = c("0.25", "0.50", "1.00")),
    Predictor_set = factor(Predictor_set, levels = c("Bulk", "FTICR", "Bulk+FTICR")),
    Feature_family = factor(
      Feature_family,
      levels = feature_family_levels
    )
  )

write_csv(section34_importance_family,
          derived_file("section34_vh0_025_050_100_importance_family.csv"))

p_section34_importance_family <- section34_importance_family %>%
  filter(family_importance > 0) %>%
  ggplot(aes(x = Vh0_display, y = family_share, fill = Feature_family)) +
  geom_col(width = 0.72, color = "white", linewidth = 0.2) +
  facet_wrap(~ Predictor_set) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = feature_family_colors, drop = FALSE) +
  labs(
    x = "Fixed Vh0",
    y = "Share of permutation importance",
    fill = "Predictor family"
  ) +
  theme_review() +
  theme(
    legend.position = "right",
    legend.box = "vertical",
    axis.text.x = element_text(size = 9)
  )

ggsave(figure_file("section34_vh0_importance_family_shift.pdf"),
       p_section34_importance_family, width = 7.4, height = 3.2)

p_section34_importance_comparison <- ggplot(
  section34_importance_comparison_top,
  aes(x = Importance, y = Feature_plot, color = Vh0_display)
) +
  geom_segment(aes(x = 0, xend = Importance, yend = Feature_plot),
               linewidth = 0.6, alpha = 0.65) +
  geom_point(size = 1.9) +
  facet_wrap(~ Predictor_set, scales = "free_y") +
  scale_y_discrete(labels = function(x) sub("^[^:]+: ", "", x)) +
  scale_color_manual(values = c(
    "0.25" = puor$dark_purple,
    "0.50" = puor$purple,
    "1.00" = puor$orange
  )) +
  labs(
    x = "Permutation importance",
    y = NULL,
    color = "Fixed Vh0"
  ) +
  theme_review() +
  theme(
    axis.text.y = element_text(size = 8),
    legend.position = "bottom"
  )

ggsave(figure_file("section34_vh0_importance_comparison.pdf"),
       p_section34_importance_comparison, width = 8.2, height = 5.2)

p_section34_importance_comparison
```

## 9. Supplement Figures: All Predictor Scenarios

This section exports the standalone supplementary sensitivity figure for all predictor scenarios.

```{r section34-supplement-all-scenarios}
section34_supp_delta_all <- kgml_competition %>%
  mutate(
    Validation = factor(
      Validation,
      levels = c("Grouped 5-fold CV", "Leave-one-site-out CV")
    ),
    Predictor_set = factor(Predictor_set, levels = predictor_set_levels),
    Vh0_label = factor(Vh0_label, levels = vh0_grid$Vh0_label),
    Line_type = if_else(Predictor_set == "Bulk+FTICR",
                        "Bulk+FTICR", "Primary predictor sets")
  )

write_csv(section34_supp_delta_all %>% select(-Line_type),
          derived_file("section34_supp_delta_rmse_all_predictor_sets.csv"))

p_section34_supp_delta_all <- ggplot(
  section34_supp_delta_all,
  aes(x = Vh0_label, y = delta_RMSE_vs_best_pure,
      color = Predictor_set, group = Predictor_set, linetype = Line_type)
) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.45) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.6) +
  facet_wrap(~ Validation, nrow = 1) +
  scale_x_discrete(labels = vh0_axis_labels) +
  scale_color_manual(values = predictor_set_colors, drop = FALSE) +
  scale_linetype_manual(
    values = c("Primary predictor sets" = "solid", "Bulk+FTICR" = "dashed"),
    guide = "none"
  ) +
  labs(
    x = "Fixed Vh0",
    y = expression(Delta * "RMSE vs best pure ML"),
    color = "Predictor set"
  ) +
  theme_review() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right"
  )

ggsave(figure_file("section34_supp_delta_rmse_all_predictor_sets.pdf"),
       p_section34_supp_delta_all, width = 8.2, height = 3.4)

p_section34_supp_delta_all
```

## 10. Pure ML Predictor-Family Diagnostic

This section computes matching predictor-family importance diagnostics for pure RF and XGBoost models.

```{r pure-ml-family-diagnostic}
fit_pure_rf_importance <- function(dat, predictors, predictor_set) {
  fixed <- mean_impute_from_train_clean(dat, dat, predictors)$train

  fit <- ranger(
    formula = reformulate(predictors, response = "logR_obs"),
    data = fixed,
    num.trees = rf_trees,
    min.node.size = 5,
    importance = "permutation",
    seed = 7
  )

  tibble(
    Predictor_set = predictor_set,
    Model = "Pure RF",
    Importance_type = "Permutation",
    Feature = names(fit$variable.importance),
    Importance = as.numeric(fit$variable.importance)
  )
}

fit_pure_xgb_importance <- function(dat, predictors, predictor_set) {
  fixed <- mean_impute_from_train_clean(dat, dat, predictors)
  scaled <- scale_from_train_clean(fixed$train, fixed$train, predictors)$train

  dtrain <- xgb.DMatrix(
    data = as.matrix(scaled[, predictors, drop = FALSE]),
    label = scaled$logR_obs
  )

  set.seed(7)
  fit <- xgb.train(
    params = list(
      objective = "reg:squarederror",
      eta = 0.04,
      max_depth = 2,
      subsample = 0.8,
      colsample_bytree = 0.8,
      min_child_weight = 3
    ),
    data = dtrain,
    nrounds = 120,
    verbose = 0
  )

  xgb.importance(model = fit, feature_names = predictors) %>%
    as_tibble() %>%
    transmute(
      Predictor_set = predictor_set,
      Model = "Pure XGBoost",
      Importance_type = "Gain",
      Feature = Feature,
      Importance = Gain
    )
}

pure_ml_importance <- predictor_spec %>%
  filter(Predictor_set %in% c("Bulk", "FTICR", "Bulk+FTICR")) %>%
  pmap_dfr(function(Predictor_set, base_predictors, include_fticr) {
    base_predictors <- unlist(base_predictors, use.names = FALSE)
    dat <- complete_for_predictor_set(
      analysis_base, base_predictors, include_fticr
    )
    predictors <- fold_predictors(dat, base_predictors, include_fticr)

    bind_rows(
      fit_pure_rf_importance(dat, predictors, Predictor_set),
      fit_pure_xgb_importance(dat, predictors, Predictor_set)
    )
  }) %>%
  mutate(
    Feature_family = classify_feature_family(Feature),
    Importance_positive = pmax(Importance, 0)
  ) %>%
  filter(Feature_family != "Other")

pure_ml_importance_family <- pure_ml_importance %>%
  group_by(Predictor_set, Model, Importance_type, Feature_family) %>%
  summarise(
    family_importance = sum(Importance_positive, na.rm = TRUE),
    n_features = dplyr::n(),
    .groups = "drop"
  ) %>%
  group_by(Predictor_set, Model, Importance_type) %>%
  mutate(
    family_share = family_importance / sum(family_importance, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    Predictor_set = factor(Predictor_set, levels = c("Bulk", "FTICR", "Bulk+FTICR")),
    Feature_family = factor(
      Feature_family,
      levels = feature_family_levels
    ),
    Model = factor(Model, levels = c("Pure RF", "Pure XGBoost"))
  )

write_csv(pure_ml_importance,
          derived_file("section34_pure_ml_predictor_importance.csv"))
write_csv(pure_ml_importance_family,
          derived_file("section34_pure_ml_importance_family.csv"))

p_pure_ml_importance_family <- pure_ml_importance_family %>%
  filter(family_importance > 0) %>%
  ggplot(aes(x = Model, y = family_share, fill = Feature_family)) +
  geom_col(width = 0.68, color = "white", linewidth = 0.2) +
  facet_wrap(~ Predictor_set) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = feature_family_colors, drop = FALSE) +
  labs(
    x = NULL,
    y = "Share of model importance",
    fill = "Predictor family"
  ) +
  theme_review() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    axis.text.x = element_text(size = 9)
  )

ggsave(figure_file("section34_pure_ml_importance_family_diagnostic.pdf"),
       p_pure_ml_importance_family, width = 7.8, height = 4.6)

p_pure_ml_importance_family
```

## 11. Predictor Manifest

This section records the predictor sets and modeling rules used by the clean validation workflow.

```{r predictor-manifest}
clean_predictor_manifest <- bind_rows(
  predictor_spec %>%
    transmute(
      Predictor_set,
      Model_class = "KGML RF residual",
      Target = "dR = logR_obs - logR_mech",
      Base_predictors = map_chr(
        base_predictors,
        ~ paste(unlist(.x, use.names = FALSE), collapse = "; ")
      ),
      FTICR_predictors = if_else(
        include_fticr,
        paste(fticr_available, collapse = "; "),
        ""
      ),
      Includes_FTICR = include_fticr,
      Rule = "Measured or direct measurement-derived predictors only"
    ),
  predictor_spec %>%
    transmute(
      Predictor_set,
      Model_class = "Pure RF/XGBoost",
      Target = "logR_obs",
      Base_predictors = map_chr(
        base_predictors,
        ~ paste(unlist(.x, use.names = FALSE), collapse = "; ")
      ),
      FTICR_predictors = if_else(
        include_fticr,
        paste(fticr_available, collapse = "; "),
        ""
      ),
      Includes_FTICR = include_fticr,
      Rule = "Same measured predictor set as paired KGML model"
    )
)

write_csv(clean_predictor_manifest,
          derived_file("clean_validation_predictor_manifest.csv"))

clean_predictor_manifest
```
