---
title: "07 Transferability Analysis"
author: "Jianqiu Zheng"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
source("R/helpers.R")
library(rsample)
library(ranger)
library(xgboost)
library(caret)
library(patchwork)

set.seed(42)
rf_trees <- 600
spatial_vh0 <- 0.50
spatial_vh0_label <- "0.50"
```

## Purpose

This workflow tests model transferability using leave-one-site-out validation.
For each held-out site, models are trained on all other sites, predicted on the
held-out site, and compared with the held-out site's nearest environmental and
geographic distance to the training domain.

The analysis focuses on Bulk and FTICR predictor sets because NPOC-only models
were already shown to have limited skill. Environmental distance is calculated
separately in each model's predictor domain, while geographic distance is
calculated from site coordinates.

## 1. Data and Predictor Sets

```{r load-data}
mech <- read_csv(derived_file("mechanistic_predictions.csv"),
                 col_types = cols(site = col_character()))

ft <- read_csv(raw_file("fticr_features.csv"),
               col_types = cols(site = col_character()))

metadata <- read_csv(raw_file("v4_WHONDRS_S19S_Metadata.csv"),
                     show_col_types = FALSE) %>%
  filter(Country == "USA") %>%
  transmute(
    site = unify_site_id(Sample_ID),
    Latitude = suppressWarnings(as.numeric(MS_Latitude_dec.deg)),
    Longitude = suppressWarnings(as.numeric(MS_Longitude_dec.deg))
  ) %>%
  filter(!is.na(site), is.finite(Latitude), is.finite(Longitude)) %>%
  distinct(site, .keep_all = TRUE)

fticr_cols <- c(
  "n_peaks", "shannon",
  "NOSC_med", "NOSC_iqr", "NOSC_mean", "NOSC_p10", "NOSC_p90",
  "NOSC_range90", "DBE_med", "DBE_mean", "DBE_iqr", "AI_med", "AI_iqr",
  "delG_med", "delG_iqr", "lambda_med", "lambda_iqr", "Cnum_med",
  "Cnum_iqr", "frac_NOSC_pos", "frac_NOSC_low", "frac_AI_high",
  "DBE_C_ratio_mean", "frac_ConHC", "frac_Lipid", "frac_AminoSugar",
  "frac_Protein", "frac_Tannin", "frac_UnsatHC", "frac_Lignin",
  "frac_Carb", "frac_Other", "ratio_Protein_Lignin",
  "ratio_Lipid_Carb", "ratio_UnsatHC_ConHC"
)

fticr_available <- intersect(fticr_cols, names(ft))

analysis_base <- mech %>%
  select(-any_of(fticr_available)) %>%
  left_join(
    ft %>% select(site, Position, all_of(fticr_available)),
    by = c("site", "Position")
  ) %>%
  left_join(metadata, by = "site") %>%
  mutate(
    log_npoc = log10(npoc),
    inv_npoc = 1 / npoc,
    logR_obs = safe_log10(respC)
  ) %>%
  filter(
    complete.cases(site, Position, respC, npoc, log_npoc, inv_npoc,
                   biomass_log10, stoichMet_donor_mean, Vh,
                   Latitude, Longitude),
    respC > 0, npoc > 0
  )

npoc_predictors <- c("npoc", "inv_npoc", "log_npoc")

bulk_predictors <- c(
  "npoc", "inv_npoc", "log_npoc",
  "biomass_log10", "delGcat_mean", "lambda_mean", "ne_mean",
  "stoichMet_donor_mean", "clay", "silt", "fine", "med", "coarse",
  "ssa_tot", "d50_mm"
)

predictor_spec <- tribble(
  ~Predictor_set, ~base_predictors, ~include_fticr,
  "Bulk", list(bulk_predictors), FALSE,
  "FTICR", list(npoc_predictors), TRUE
)

analysis_base %>%
  summarise(samples = n(), sites = n_distinct(site))
```

## 2. Modeling Helpers

```{r helpers}
make_loso_folds_clean <- function(dat) {
  sites <- sort(unique(dat$site))
  splits <- lapply(sites, function(s) {
    make_splits(
      list(analysis = which(dat$site != s), assessment = which(dat$site == s)),
      dat
    )
  })

  tibble(splits = splits, id = paste0("LOSO_", sites), holdout_site = sites)
}

fold_id_clean <- function(folds, i) {
  if ("id" %in% names(folds)) as.character(folds$id[[i]]) else paste0("Fold", i)
}

mean_impute_from_train_clean <- function(train_df, test_df, predictors) {
  train_out <- train_df
  test_out <- test_df

  for (nm in predictors) {
    fill <- mean(train_out[[nm]], na.rm = TRUE)
    if (!is.finite(fill)) fill <- 0
    train_out[[nm]][!is.finite(train_out[[nm]]) | is.na(train_out[[nm]])] <- fill
    test_out[[nm]][!is.finite(test_out[[nm]]) | is.na(test_out[[nm]])] <- fill
  }

  list(train = train_out, test = test_out)
}

scale_from_train_clean <- function(train_df, test_df, predictors) {
  train_out <- train_df
  test_out <- test_df

  for (nm in predictors) {
    center <- mean(train_out[[nm]], na.rm = TRUE)
    spread <- sd(train_out[[nm]], na.rm = TRUE)
    if (!is.finite(center)) center <- 0
    if (!is.finite(spread) || spread == 0) spread <- 1
    train_out[[nm]] <- (train_out[[nm]] - center) / spread
    test_out[[nm]] <- (test_out[[nm]] - center) / spread
  }

  list(train = train_out, test = test_out)
}

prune_fticr_from_train <- function(train_df, candidates, cutoff = 0.80) {
  candidates <- candidates[candidates %in% names(train_df)]
  candidates <- candidates[vapply(train_df[candidates], is.numeric, logical(1))]

  if (length(candidates) < 2) return(candidates)

  complete_counts <- vapply(train_df[candidates], function(x) {
    sum(is.finite(x))
  }, numeric(1))
  candidates <- candidates[complete_counts >= 5]
  if (length(candidates) < 2) return(candidates)

  corr_mat <- suppressWarnings(cor(train_df[, candidates, drop = FALSE],
                                   use = "pairwise.complete.obs"))
  corr_mat[!is.finite(corr_mat)] <- 0
  high_corr <- caret::findCorrelation(corr_mat, cutoff = cutoff)
  if (length(high_corr) > 0) candidates[-high_corr] else candidates
}

add_fixed_vh0_mechanistic <- function(dat, fixed_vh0) {
  dat %>%
    mutate(
      Vh0 = fixed_vh0,
      R_mech = exp(-abs(stoichMet_donor_mean) / (fixed_vh0 * npoc)) *
        biomass_log10,
      logR_mech = safe_log10(R_mech),
      dR = logR_obs - logR_mech
    )
}

complete_for_predictor_set <- function(dat, base_predictors, include_fticr) {
  required <- base_predictors
  if (include_fticr) required <- c(required, fticr_available)

  dat %>%
    filter(if_all(all_of(required), is.finite))
}

fold_predictors <- function(train_df, base_predictors, include_fticr) {
  if (!include_fticr) return(base_predictors)
  unique(c(base_predictors, prune_fticr_from_train(train_df, fticr_available)))
}

fit_rf_predict <- function(train_df, test_df, response, predictors) {
  predictors <- predictors[predictors %in% names(train_df)]
  fixed <- mean_impute_from_train_clean(train_df, test_df, predictors)

  fit <- ranger(
    formula = reformulate(predictors, response = response),
    data = fixed$train,
    num.trees = rf_trees,
    min.node.size = 5,
    importance = "permutation",
    seed = 7
  )

  predict(fit, data = fixed$test)$predictions
}

fit_xgb_predict <- function(train_df, test_df, response, predictors) {
  predictors <- predictors[predictors %in% names(train_df)]
  fixed <- mean_impute_from_train_clean(train_df, test_df, predictors)
  scaled <- scale_from_train_clean(fixed$train, fixed$test, predictors)

  dtrain <- xgb.DMatrix(
    data = as.matrix(scaled$train[, predictors, drop = FALSE]),
    label = scaled$train[[response]]
  )
  dtest <- xgb.DMatrix(data = as.matrix(scaled$test[, predictors, drop = FALSE]))

  set.seed(7)
  fit <- xgb.train(
    params = list(
      objective = "reg:squarederror",
      eta = 0.04,
      max_depth = 2,
      subsample = 0.8,
      colsample_bytree = 0.8,
      min_child_weight = 3
    ),
    data = dtrain,
    nrounds = 120,
    verbose = 0
  )

  as.numeric(predict(fit, dtest))
}
```

## 3. Transfer Distance Metrics

```{r environmental-distance-helper}
calc_nearest_environmental_distance <- function(train_df, test_df, predictors) {
  fixed <- mean_impute_from_train_clean(train_df, test_df, predictors)

  train_site_env <- fixed$train %>%
    group_by(site) %>%
    summarise(across(all_of(predictors), ~ mean(.x, na.rm = TRUE)),
              .groups = "drop")

  test_site_env <- fixed$test %>%
    group_by(site) %>%
    summarise(across(all_of(predictors), ~ mean(.x, na.rm = TRUE)),
              .groups = "drop")

  centers <- map_dbl(train_site_env[predictors], ~ mean(.x, na.rm = TRUE))
  spreads <- map_dbl(train_site_env[predictors], ~ sd(.x, na.rm = TRUE))
  spreads[!is.finite(spreads) | spreads == 0] <- 1

  train_scaled <- sweep(as.matrix(train_site_env[predictors]), 2, centers, "-")
  train_scaled <- sweep(train_scaled, 2, spreads, "/")
  test_scaled <- sweep(as.matrix(test_site_env[predictors]), 2, centers, "-")
  test_scaled <- sweep(test_scaled, 2, spreads, "/")

  env_dist <- map_dbl(seq_len(nrow(test_scaled)), function(j) {
    d <- sqrt(rowSums((sweep(train_scaled, 2, test_scaled[j, ], "-"))^2))
    min(d, na.rm = TRUE)
  })

  tibble(
    site = test_site_env$site,
    environmental_dissimilarity = env_dist,
    n_train_sites = nrow(train_site_env),
    n_environmental_predictors = length(predictors)
  )
}

haversine_km <- function(lon1, lat1, lon2, lat2) {
  radius_km <- 6371
  to_rad <- pi / 180
  lon1 <- lon1 * to_rad
  lat1 <- lat1 * to_rad
  lon2 <- lon2 * to_rad
  lat2 <- lat2 * to_rad

  dlon <- lon2 - lon1
  dlat <- lat2 - lat1
  a <- sin(dlat / 2)^2 + cos(lat1) * cos(lat2) * sin(dlon / 2)^2
  2 * radius_km * atan2(sqrt(a), sqrt(1 - a))
}

calc_nearest_geographic_distance <- function(train_df, test_df) {
  train_sites <- train_df %>%
    group_by(site) %>%
    summarise(
      Latitude = mean(Latitude, na.rm = TRUE),
      Longitude = mean(Longitude, na.rm = TRUE),
      .groups = "drop"
    )

  test_sites <- test_df %>%
    group_by(site) %>%
    summarise(
      Latitude = mean(Latitude, na.rm = TRUE),
      Longitude = mean(Longitude, na.rm = TRUE),
      .groups = "drop"
    )

  nearest_km <- map_dbl(seq_len(nrow(test_sites)), function(i) {
    d <- haversine_km(
      lon1 = test_sites$Longitude[[i]],
      lat1 = test_sites$Latitude[[i]],
      lon2 = train_sites$Longitude,
      lat2 = train_sites$Latitude
    )
    min(d, na.rm = TRUE)
  })

  test_sites %>%
    transmute(
      site,
      Latitude,
      Longitude,
      nearest_training_distance_km = nearest_km,
      n_train_sites_geo = nrow(train_sites)
    )
}
```

## 4. Run LOSO Transfer Test

```{r run-loso-transfer}
run_loso_for_predictor_set <- function(spec_row) {
  predictor_set <- spec_row$Predictor_set
  base_predictors <- unlist(spec_row$base_predictors[[1]], use.names = FALSE)
  include_fticr <- spec_row$include_fticr

  dat <- complete_for_predictor_set(
    analysis_base, base_predictors, include_fticr
  )

  folds <- make_loso_folds_clean(dat)
  out <- vector("list", nrow(folds))
  dist_out <- vector("list", nrow(folds))
  geo_out <- vector("list", nrow(folds))

  for (i in seq_len(nrow(folds))) {
    train_raw <- analysis(folds$splits[[i]])
    test_raw <- assessment(folds$splits[[i]])
    fold_id <- fold_id_clean(folds, i)
    holdout_site <- as.character(folds$holdout_site[[i]])

    train_vh <- train_raw %>% add_fixed_vh0_mechanistic(spatial_vh0)
    test_vh <- test_raw %>% add_fixed_vh0_mechanistic(spatial_vh0)
    predictors <- fold_predictors(train_vh, base_predictors, include_fticr)

    pure_rf <- fit_rf_predict(train_raw, test_raw, "logR_obs", predictors)
    pure_xgb <- fit_xgb_predict(train_raw, test_raw, "logR_obs", predictors)
    kgml_dR <- fit_rf_predict(train_vh, test_vh, "dR", predictors)

    out[[i]] <- bind_rows(
      test_raw %>%
        transmute(
          site, Position, fold_id = fold_id, holdout_site = holdout_site,
          Predictor_set = predictor_set,
          Model = paste("Pure RF", predictor_set),
          Vh0_label = NA_character_,
          Vh0 = NA_real_,
          logR_obs,
          logR_pred = pure_rf
        ),
      test_raw %>%
        transmute(
          site, Position, fold_id = fold_id, holdout_site = holdout_site,
          Predictor_set = predictor_set,
          Model = paste("Pure XGBoost", predictor_set),
          Vh0_label = NA_character_,
          Vh0 = NA_real_,
          logR_obs,
          logR_pred = pure_xgb
        ),
      test_vh %>%
        transmute(
          site, Position, fold_id = fold_id, holdout_site = holdout_site,
          Predictor_set = predictor_set,
          Model = paste("KGML RF", predictor_set),
          Vh0_label = spatial_vh0_label,
          Vh0 = spatial_vh0,
          logR_obs,
          logR_pred = logR_mech + kgml_dR
        )
    )

    dist_out[[i]] <- calc_nearest_environmental_distance(
      train_vh, test_vh, predictors
    ) %>%
      mutate(
        fold_id = fold_id,
        holdout_site = holdout_site,
        Predictor_set = predictor_set
      )

    geo_out[[i]] <- calc_nearest_geographic_distance(train_vh, test_vh) %>%
      mutate(
        fold_id = fold_id,
        holdout_site = holdout_site,
        Predictor_set = predictor_set
      )
  }

  list(
    predictions = bind_rows(out),
    distances = bind_rows(dist_out),
    geographic_distances = bind_rows(geo_out)
  )
}

loso_results <- map(
  seq_len(nrow(predictor_spec)),
  ~ run_loso_for_predictor_set(predictor_spec[.x, ])
)

envdist_loso_predictions <- map_dfr(loso_results, "predictions") %>%
  mutate(
    R_obs = 10^logR_obs,
    R_pred = 10^logR_pred,
    residual_log = logR_obs - logR_pred
  )

envdist_loso_distances <- map_dfr(loso_results, "distances")
envdist_loso_geographic_distances <- map_dfr(loso_results, "geographic_distances")

write_csv(envdist_loso_predictions,
          derived_file("transferability_loso_predictions.csv"))
write_csv(envdist_loso_distances,
          derived_file("transferability_environmental_dissimilarity.csv"))
write_csv(envdist_loso_geographic_distances,
          derived_file("transferability_geographic_distance.csv"))

envdist_loso_stats <- envdist_loso_predictions %>%
  group_by(Predictor_set, Model, Vh0_label, Vh0) %>%
  group_modify(~ model_stats(.x$logR_obs, .x$logR_pred)) %>%
  ungroup() %>%
  select(-R2) %>%
  arrange(Predictor_set, Model)

write_csv(envdist_loso_stats,
          derived_file("transferability_loso_model_stats.csv"))

envdist_loso_stats
```

## 5. LOSO Model Performance Summary

```{r loso-performance-lollipop}
envdist_best_pure <- envdist_loso_stats %>%
  filter(str_detect(Model, "^Pure ")) %>%
  group_by(Predictor_set) %>%
  slice_min(RMSE, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(
    Predictor_set,
    best_pure_model = Model,
    best_pure_RMSE = RMSE,
    best_pure_CCC = CCC
  )

transferability_model_summary <- bind_rows(
  envdist_loso_stats %>%
    filter(str_detect(Model, "^KGML ")) %>%
    mutate(Model_display = "KGML RF"),
  envdist_loso_stats %>%
    inner_join(envdist_best_pure,
               by = c("Predictor_set", "Model" = "best_pure_model")) %>%
    mutate(Model_display = "Best pure RF/XGBoost")
) %>%
  mutate(
    Predictor_set = factor(Predictor_set, levels = c("Bulk", "FTICR")),
    Model_display = factor(
      Model_display,
      levels = c("Best pure RF/XGBoost", "KGML RF")
    )
  )

transferability_metric_long <- transferability_model_summary %>%
  select(Predictor_set, Model, Model_display, RMSE, CCC) %>%
  pivot_longer(c(RMSE, CCC), names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = factor(Metric, levels = c("RMSE", "CCC")))

write_csv(transferability_model_summary,
          derived_file("transferability_loso_model_summary.csv"))
write_csv(transferability_metric_long,
          derived_file("transferability_loso_metric_long.csv"))

p_loso_performance <- ggplot(
  transferability_metric_long,
  aes(x = Value, y = Predictor_set, color = Model_display)
) +
  geom_segment(
    aes(x = 0, xend = Value, yend = Predictor_set),
    linewidth = 0.65,
    alpha = 0.7,
    position = position_dodge(width = 0.55)
  ) +
  geom_point(position = position_dodge(width = 0.55), size = 2.4) +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  scale_color_manual(values = c(
    "KGML RF" = puor$purple,
    "Best pure RF/XGBoost" = puor$orange
  )) +
  labs(
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme_review() +
  theme(legend.position = "right")

ggsave(figure_file("transferability_loso_performance_lollipop.pdf"),
       p_loso_performance, width = 7.8, height = 2.8)

p_loso_performance
```

## 6. Error Versus Transfer Distance

```{r error-distance-analysis}
envdist_site_errors <- envdist_loso_predictions %>%
  mutate(abs_error = abs(logR_obs - logR_pred)) %>%
  group_by(Predictor_set, Model, Vh0_label, Vh0, fold_id, holdout_site, site) %>%
  summarise(
    n_samples = n(),
    mean_abs_error = mean(abs_error, na.rm = TRUE),
    RMSE = sqrt(mean((logR_obs - logR_pred)^2, na.rm = TRUE)),
    Bias = mean(logR_pred - logR_obs, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    envdist_loso_distances,
    by = c("Predictor_set", "fold_id", "holdout_site", "site")
  ) %>%
  left_join(
    envdist_loso_geographic_distances,
    by = c("Predictor_set", "fold_id", "holdout_site", "site")
  ) %>%
  left_join(envdist_best_pure, by = "Predictor_set") %>%
  mutate(
    Model_display = case_when(
      str_detect(Model, "^KGML ") ~ "KGML RF",
      Model == best_pure_model ~ "Best pure RF/XGBoost",
      TRUE ~ NA_character_
    ),
    Predictor_set = factor(Predictor_set, levels = c("Bulk", "FTICR"))
  )

envdist_plot_data <- envdist_site_errors %>%
  filter(!is.na(Model_display))

envdist_error_summary <- envdist_plot_data %>%
  group_by(Predictor_set, Model, Model_display) %>%
  summarise(
    n_sites = n_distinct(site),
    spearman_rho = suppressWarnings(
      cor(mean_abs_error, environmental_dissimilarity,
          use = "complete.obs", method = "spearman")
    ),
    linear_slope = coef(lm(mean_abs_error ~ environmental_dissimilarity))[2],
    .groups = "drop"
  ) %>%
  arrange(Predictor_set, Model_display)

geo_error_summary <- envdist_plot_data %>%
  group_by(Predictor_set, Model, Model_display) %>%
  summarise(
    n_sites = n_distinct(site),
    spearman_rho = suppressWarnings(
      cor(mean_abs_error, nearest_training_distance_km,
          use = "complete.obs", method = "spearman")
    ),
    linear_slope = coef(lm(mean_abs_error ~ nearest_training_distance_km))[2],
    .groups = "drop"
  ) %>%
  arrange(Predictor_set, Model_display)

transfer_distance_summary <- bind_rows(
  envdist_error_summary %>%
    mutate(Distance_type = "Environmental dissimilarity"),
  geo_error_summary %>%
    mutate(Distance_type = "Geographic distance")
) %>%
  mutate(
    Distance_type = factor(
      Distance_type,
      levels = c("Environmental dissimilarity", "Geographic distance")
    )
  )

envdist_competition <- envdist_loso_stats %>%
  filter(str_detect(Model, "^KGML ")) %>%
  left_join(envdist_best_pure, by = "Predictor_set") %>%
  mutate(
    delta_RMSE_vs_best_pure = RMSE - best_pure_RMSE,
    delta_CCC_vs_best_pure = CCC - best_pure_CCC,
    beats_best_pure = delta_RMSE_vs_best_pure < 0
  )

write_csv(envdist_site_errors,
          derived_file("transferability_loso_site_errors.csv"))
write_csv(envdist_error_summary,
          derived_file("transferability_error_dissimilarity_summary.csv"))
write_csv(geo_error_summary,
          derived_file("transferability_error_geographic_distance_summary.csv"))
write_csv(transfer_distance_summary,
          derived_file("transferability_distance_summary.csv"))
write_csv(envdist_competition,
          derived_file("transferability_kgml_vs_best_pure.csv"))

p_error_distance <- ggplot(
  envdist_plot_data,
  aes(x = environmental_dissimilarity, y = mean_abs_error,
      color = Model_display)
) +
  geom_point(alpha = 0.72, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.85) +
  facet_wrap(~ Predictor_set, scales = "free_x", nrow = 1) +
  scale_color_manual(values = c(
    "KGML RF" = puor$purple,
    "Best pure RF/XGBoost" = puor$orange
  )) +
  labs(
    x = "Nearest-training environmental dissimilarity",
    y = "Held-out site mean absolute error",
    color = NULL
  ) +
  theme_review() +
  theme(legend.position = "right")
ggsave(figure_file("environment_dissimilarity_scatter.pdf"),
       p_error_distance, width = 7.8, height = 3)


p_error_geo_distance <- ggplot(
  envdist_plot_data,
  aes(x = nearest_training_distance_km, y = mean_abs_error,
      color = Model_display)
) +
  geom_point(alpha = 0.72, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.85) +
  facet_wrap(~ Predictor_set, scales = "free_x", nrow = 1) +
  scale_color_manual(values = c(
    "KGML RF" = puor$purple,
    "Best pure RF/XGBoost" = puor$orange
  )) +
  labs(
    x = "Nearest-training geographic distance (km)",
    y = "Held-out site mean absolute error",
    color = NULL
  ) +
  theme_review() +
  theme(legend.position = "right")
ggsave(figure_file("geographic_distance_scatter.pdf"),
       p_error_geo_distance, width = 7.8, height = 3)



p_error_rho <- ggplot(
  transfer_distance_summary,
  aes(x = Predictor_set, y = spearman_rho, fill = Model_display)
) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.35) +
  geom_col(position = position_dodge(width = 0.68),
           width = 0.58, color = "black", linewidth = 0.2) +
  facet_wrap(~ Distance_type, nrow = 1) +
  scale_fill_manual(values = c(
    "KGML RF" = puor$purple,
    "Best pure RF/XGBoost" = puor$orange
  )) +
  labs(
    x = NULL,
    y = "Error-distance correlation",
    fill = NULL
  ) +
  theme_review() +
  theme(legend.position = "right")

ggsave(figure_file("transferability_distance_rho_summary.pdf"),
       p_error_rho, width = 7.8, height = 3)

p_section35_transferability <- p_loso_performance /
  p_error_distance /
  p_error_geo_distance /
  p_error_rho +
  plot_layout(heights = c(0.9, 1, 1, 0.9)) +
  plot_annotation(tag_levels = "a")

ggsave(figure_file("section35_transferability_summary.pdf"),
       p_section35_transferability, width = 8.0, height = 10.2)

p_section35_transferability

```
